---
title: "Pre-registered report: Mapping Synesthetic Sequences in Space: Improving Consistency Test by Harnessing Cartography Tools."
shorttitle: "Space-Sequence Synesthesia: Optimizing Consistency Tests"
author:
  - name: Rémy Lachelin
    corresponding: true
    orcid: 0000-0002-8485-7153
    role: 
      - Conceptualization
      - Methodology
      - Formal Analysis
      - Data Curation
      - Writing - Original Draft
    email: remy.lachelin@fernuni.ch
    affiliations:
      - name: UniDistance Suisse
        department: Psychology
        address: Schinerstrasse 18
        city: Brig-Glis
        region: Valais
        postal-code: 3900  
  - name: Chhavi Sachdeva
    corresponding: false
    orcid: 0000-0002-0074-4371
    role: 
      - Conceptualization
      - Methodology
      - Writing - Review & Editing
    email: chhavi.sachdeva@unidistance.ch
    affiliations:
      - name: UniDistance Suisse
        department: Psychology
        address: Schinerstrasse 18
        city: Brig-Glis
        region: Valais
        postal-code: 3900     
  - name: Nicolas Rothen
    corresponding: false
    orcid: 0000-0002-8874-8341
    role: 
      - Conceptualization
      - Writing - Review & Editing
      - Supervision
      - Project Administration
      - Founding Acquisition
    email: nicolas.rothen@fernuni.ch
    affiliations:
      - name: UniDistance Suisse
        department: Psychology
        address: Schinerstrasse 18
        city: Brig-Glis
        region: Valais
        postal-code: 3900   
abstract: "Sequence-space synesthesia is a perceptual condition in which ordinal sequences (e.g., numbers, days of the week, or months of the year) are experienced as occupying specific spatial locations. This perceptual condition is identified using self-reports (e.g. questionnaires) and behavioural consistency tasks. Existing consistency tests measure the spatial distance between responses to repeated stimuli, but they ignore the ordinal and geometric features of spatial-forms. This preregistered study consists of a present and future analyses. In the present analyses, we optimise SSS diagnostics from consistency test by extracting novel geometric features at the spatial-form level. To do this, we first evaluate classification performances of old and new features in a large (N = 685) aggregated sample from four available and independent datasets. We then harness a geographic toolbox to extract metrics from the spatial-forms level. Finally, receiver operating characteristic (ROC) analyses are conducted to assess the discriminative performances of each method. The results show that permuted topological validity of spatial-forms and the perimeter between stimuli performs equally well in detecting SSS. In a future study, we will examine predictive validity of the selected methods on an independent dataset that has yet to be collected. Findings from the present study highlight the relevance of topological principles for sequence-space representations and aims to optimise the ability of consistency tests to detect SSS."
keywords: [Sequence space synaesthesia, Synaesthesia/synesthesia, consistency test, space, time, numbers]
author-note:
  disclosures:
    conflict of interest: The author has no conflict of interest to declare.
bibliography: references.bib
numbered-lines: true
format:
  apaquarto-pdf:
    number-sections: true
    floatsintext: false
    documentmode: man
    a4paper: true
    donotrepeattitle: true
    keep-tex: true
#  apaquarto-html:
#    crossrefs-hover: true
#    floatsintext: true
#    hypothesis: true
# project:
#  type: website
#  output-dir: docs
#  apaquarto-typst: 
#    fig-format: png
execute:
  warning: false
  message: false
  cache: true
editor: 
  markdown: 
    wrap: 72
---

```{r, libraries, message=FALSE, warning=FALSE}
# To reproduce the manuscript:
# 1. Download all the github repository (including data [.csv] and bibliography [.bib])
# 2. Requires Quarto (RStudio) + apa-quarto template (https://github.com/wjschne/apaquarto).
#     To install apa-quarto template:
#     quarto::quarto_use_template("wjschne/apaquarto")
#     (Or in the terminal: quarto use template wjschne/apaquarto )
# 3. Requires the following packages to be installed 
#     (i.e. install.package("package_name"))
# 4. Click on render (optimized for .pdf format (via LateX)).

# Data import
library(readr)
library(readxl)
# Data wrangling:
library(tidyr)
library(dplyr)
# Render Formatting
# library(papaja)
library(RColorBrewer)
# library(kableExtra)
library(pakret) # R package in text citation
# Plots
library(ggplot2)
library(ggridges)
library(ggalluvial)
library(ggVennDiagram)
library(ggridges)
library(ggforce)
library(cowplot) # for multiplot in 1 figure
library(corrplot)
# geometry/geography feature package
library(sf)
# for ROC analyses:
library(pROC) # See https://www.r-bloggers.com/2019/02/some-r-packages-for-roc-curves/

# Make shure environment is empty
rm(list = ls())
```

# Introduction {#sec-introduction}

Sequence-Space Synaesthesia (SSS) or visuo-spatial forms is the
phenomenon where people visualize ordered sequences in particular
spatial positions. For example, numbers, weekdays or months (synesthetic
*inducers*) are represented as arranged into specific spatial positions
in space (synesthetic *concurrent*). The synaesthetic spatial-forms can
be complex and have variable geometric forms. These geometric forms
might be different for each category (i.e. number-forms weekdays and
months), with forms such as ellipses, zig-zags or curbed lines
[@galton1880,@flournoy1893]. However, spaital-forms of individuals with
SSS are consistent and automatic. Methods to classify SSS and controls
using consistency are however heterogeneous. The aim of the present
study is to optimize consistency methods to classify SSS and controls.

SSS frequently co-occurs with other subtypes of synaesthesia and as for
other subtypes of synaesthesia it’s idiosyncratic. Colour-grapheme
synaesthesia, where a grapheme *inducer* triggers a *concurrent* colour
(i.e. “A” seen as red), co-occurs at 71 % to 76 % with SSS [@sagiv2006].
Nevertheless, across different synaesthetic subtypes, self-reported SSS
tend to cluster together [@ward2022], indicating a degree of internal
homogeneity and justifying to study it as a distinct phenomenon.
Spatial-forms are idiosyncratic, which results in considerable
heterogeneity in how this phenomenon is manifested across individuals.
One source of heterogeneity is given by dimensionality: some SSS
experiences can involve three-dimensional (3D) and two-dimensional (2D)
arrangements [@eagleman2009a; @price2013]. Another source of
heterogeneity is the reference frame, for example, the spatial-forms
might take place in an external space around the body (*i.e.* projector)
or in an egocentric internal space (*i.e.* associator) [@dixon2004;
@smilek2007]. Further variability can be explained by temporal-spatial
properties and mental-manipulation of spatial-forms such as “zooming” in
and out or rotating or shifting perspectives at will [@gould2014;
@jarick2009] Lastly, the shape, complexity and layout of the
spatial-forms are also heterogeneous such as forming for example ovals,
lines, zig-zags or loops. With some recurring shapes being more frequent
across SSS, such as ovals for months [@eagleman2009a]
<!--# I think with this paragraph, it’s difficult to see how it is connected to your research question (especially the bit about co-occurence with other types of synesthesia). I wonder if you need some sentences that connect the premise of this paragraph back to the main research question. -->.

Despite the after mentioned heterogeneities, SSS is also
phenomenological characterized
<!--# I think these five characteristics are quite well-established in synesthesia, right? So, I wonder if it would better to say generally that there are five main characteristics for synesthesia. -->.
There might be five main characteristics for synaesthesia in general and
specifically SSS [@seron1992; @deroy2013]. *Automaticity*: the inducer
automatically triggers the concurrent. *Unidirectionality*: while the
*inducer* triggers the concurrent, the concurrent does not trigger the
inducer. *Developmentally*: the experience was already present during
childhood. *Consciousness*: The concurrent is consciously perceived.
*Consistency*: the inducer-concurrent pair remains stable in time within
subject. While some of these characteristics can be captured with
self-reported questionnaire (i.e. for development and consciousness),
other can be quantified more objectively using behavioural tasks such as
consistency tests
<!--# I think consistency is the only one that is captured using consistency tests. The others are all captured using self-report (atleast so far). -->[@baron-cohen1993].

## Consistency tests {#sec-consistency-tests}

Individuals with SSS are usually identified using questionnaires and
consistency tests. . However, stimulus level consistency can also be
achieved by responding to the same position for all stimuli.
Furthermore, it might favour linear over circular spatial-forms and is
not informed about the ordinality between stimuli (e.g. 1 `→` 2 `→` 3,
etc.). Here, we aim at optimizing the identification of SSS from
consistency tasks using geometrical characteristics of spatial-forms
from ordered stimuli or inducers.

Consistency tests are extracted form behavioural tasks used to evaluate
how stable synesthetic experiences are over time [@rothen2013a;
@rothen2016]. They do this by presenting the same stimuli repeatedly and
assessing how similar participant's responses are across repetitions
(i.e. *inducer*-*concurrent* associations). An individual is identified
as SSS based on the variability at the stimulus level between repeated
responses: less variable responses being an indicator of the consistent
response characteristic of SSS. Consistency tests have proven effective
for colour-grapheme synaesthesia. Measures of individual consistency can
be derived using colour pickers while repeatedly presenting the same
inducer (i.e. “A”). The quantification of the distance between
concurrents from the same inducers is used to discriminate consistent
synaesthetes from controls [@rothen2013a]. A similar rationale to that
used for colour-grapheme synaesthesia has been applied to characterize
SSS from consistency tasks.

In the SSS consistency task, participants complete multiple rounds of
the same stimuli (e.g., digits, days of the week, and months of the
year) and are asked to indicate a spatial location for each stimulus by
clicking on the computer screen. @brang2010 for example evaluated
consistency as the distance between repeated responses to the same
inducer or stimuli (e.g. January and January) relative to adjacent
stimuli (e.g. February). A response was defined as consistent if it fell
within 1.96 *z-scores*. However, this criterion was noted to be
potentially too conservatory, as it identified synaesthesia in 4 of 81
self-reported synaesthetes. Geometrical features such as area and
perimeter between the coordinates of repeated inducers have further been
used as a measure of the distance between same inducers, hence as a
metric for consistency. Less consistent responses lead to smaller
triangle perimeters and area [@rothen2016]. A cut-off corresponding to
less than 0.203 % average responses area in proportion to the total
screen area to classify as SSS has been suggested [see also @ward2018]
<!--# I am unsure as to what this sentence means. -->.

## Limitations of current methods of consistency tests

However, several general caveats of consistency tests using concurrent
or stimulus level metrics have been identified. Some tests may favour
certain spatial-forms [@ward2018] in particular those with linear or
highly regular spatial layouts, i.e. elliptical patterns for
months[@flournoy1893; @eagleman2009a; @brang2010]. Another concern is
that participants that respond with the same position, for example when
not knowing the response, will have artificially high consistencies with
those metrics [@rothen2016]. This has led to alternative approaches such
as to add the standard deviation of responses, questionnaire cut-off
[@ward2018] or permutation-based comparison of individual responses to
chance levels for colour-grapheme [@root2021] and SSS [@ward2022a].
Finally, these consistency tests evaluate the variability of responses
to the same stimuli, therefore the ordinality between stimuli is not
taken into account. Synaesthetic spatial-forms follow ordinal rules
(e.g. Monday `→` Tuesday `→` Wednesday). In the domain of numerical
cognition, ordinality has been identified as an important information
when processing sequences [@lyons2013]. Some studies of SSS have
systematically investigated ordinality by investigating metrics of
adjacent inducer-concurrent pairs such as distances [@brang2010] or
angles [@eagleman2009a]. However, ordinality of the concurrent
spatial-form might be particularly relevant considering all the stimuli
of a certain category (i.e. weekdays or months). The spatial-forms of
the concurrents are often spatially structured into configuration such
as lines or polygons that may follow distinctive geometrical rules. For
example linear layouts have been described in early accounts of number
forms [@flournoy1893; @galton1880].

## Present study

The goal of this registered report is to compare different consistency
tests on their classification performances of individuals with SSS and
controls doing the consistency task using Receiver Operator
Characteristics (ROC) analyses. In the present *Phase I*, we merge four
previously available datasets using the same task on both SSS and
control groups [@ward2022a]. First, we aimed to reproduce established
consistency test methods based on stimulus-level consistency metrics,
such as area and perimeter. Second, we explored a new approach comparing
geometrical features across repetitions of the same categories (weekday,
month and numbers), such as segments and polygons. These novel tests are
designed to take advantage of the sequentiality or ordinality between
the responses and the geometrical properties at the spatial-form level,
see @fig-myplot00. All the consistency tests are then compared using ROC
analyses on their correct classification performances of self-reported
SSS and control groups.

In a future study, we will assess whether the tests identified in the
present study generalize and are validated on an independent dataset
that is yet to be acquired (@sachdeva2024, Stage 1 in-principle
acceptance, see also <https://osf.io/6wn4m>).

```{r Load data}
# In the following I upload and merge the data from Ward, Rothen and Van Peters. Data is stored into a full dataset ds (i.e. 1 row per trial) and a dataset per participant ds_Quest (i.e. 1 row per participant).

###  Ward Data
ds_syn       <- read_excel("Data/raw_synaesthetes_consistency_anon.xlsx")
ds_syn$group <- "Syn"
ds_ctl       <- read_excel("Data/raw_controls_consistency_anon.xlsx")
ds_ctl$group <- "Ctl"

ds_Q_syn       <- read_excel("Data/raw_synaesthetes_questionnaire_anon.xlsx")
ds_Q_syn$group <- "Syn"
ds_Q_ctl       <- read_excel("Data/raw_controls_questionnaire_anon.xlsx")
ds_Q_ctl$group <- "Ctl"

# Merge wards datafiles:
ds <- merge(ds_syn,ds_ctl, all = TRUE)
ds_Q <- merge(ds_Q_syn,ds_Q_ctl, all = TRUE)

# Ward only uses those who completed the Questionnaire (i.e. N = 215+252 = 467)
ds <- ds %>% 
  filter(session_id %in% unique(ds_Q$session_id))
 
### Rothen Data

ds_Rothen <- read.csv("Data/rawdata.txt", sep="")

### Rename variables to match datasets

# sum(ds_Q$consistency_score != ds_Q$consistency) # Duplicate variable
ds_Q$consistency <- NULL
ds_Q$...36 <- NULL
ds_Q$...37 <- NULL
ds_Q$mean_simulation_Z <- NULL
ds_Q$SD_simulation <- NULL
ds_Q$`z-score`  <- NULL

rm(ds_syn,ds_ctl,ds_Q_syn,ds_Q_ctl)

ds$ID <- ds$session_id
ds_Q$ID <- ds_Q$session_id

ds_Q$dataSource <- "Ward"
ds$dataSource   <- "Ward"

# From Rothen:
names(ds_Rothen)[names(ds_Rothen) == "Group"] <- "group"
ds_Rothen$group <- as.factor(ds_Rothen$group)
levels(ds_Rothen$group) <- c("Ctl","Syn")
names(ds_Rothen)[names(ds_Rothen) == "Inducer"] <- "stimulus"
names(ds_Rothen)[names(ds_Rothen) == "X"] <- "x"
names(ds_Rothen)[names(ds_Rothen) == "Y"] <- "y"
ds_Rothen$SynQuest <- ds_Rothen$group == "Syn"

ds_Rothen$dataSource <- "Rothen"

# From the paper (all the same since lab based):
ds_Rothen$width  <- 1024
ds_Rothen$height <- 768
```

```{r Merge1 Rothen and Ward data}
## 0.2 Merge data:
# remove non matching colnames: 
ColNames_ds <- colnames(ds_Rothen)[colnames(ds_Rothen) %in% colnames(ds)]

ds <- ds %>%
  select(all_of(ColNames_ds))
ds_Rothen <- ds_Rothen %>%
  select(all_of(ColNames_ds))

# Merge Data:
ds   <- merge(ds,ds_Rothen, all = TRUE)

# Because there is no questionnaire in Nicola's data:
ds$SynQuest[ds$dataSource == "Rothen"] = "NaN"

# Questionnaire:
ID <- unique((ds_Rothen$ID))
ds_Q_Rothen <- as.data.frame(ID)
ds_Q_Rothen$dataSource <- "Rothen"
ds_Q_Rothen <- merge(ds_Q_Rothen, ds_Rothen %>% group_by(ID) %>% select(ID, dataSource, group) %>% filter(row_number() == 1), by = c("ID","dataSource"))

# Append rows:
ds_Q$ID <- as.character(ds_Q$ID)
ds_Q <-  bind_rows(ds_Q,ds_Q_Rothen)

# Clear up:
rm(ds_Q_Rothen, ds_Rothen)

## 0.3 Wrangle dataset
# Add Condition, i.e. stim type:
ds$Cond <- NaN
ds$Cond[ds$stimulus %in% c("1","2","3","4","5","6","7","8","9","0")] <- "number"
ds$Cond[ds$stimulus %in% c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")] <- "weekday"
ds$Cond[ds$stimulus %in% c("January", "February", "March", "April", "May","June","July","August","September","October","November","December")] <- "month"
```

```{r MergeVanPeters}
### ADD Van Petersen Cortex data
filedir <- "Data/vanPeterPerID/"
fn <- paste0(filedir,dir(filedir))

ds_PeterCor <- read_excel(fn[1],
                                col_types = c("text", "numeric", "text", 
                                              "text", "numeric", "numeric", "numeric", 
                                              "numeric", "numeric", "numeric", 
                                              "numeric", "numeric", "text", "text", 
                                              "text", "numeric", "numeric", "numeric"))

for(i in 2:length(fn)){
  ds_i <- read_excel(fn[i],
                     col_types = c("text", "numeric", "text", 
                                   "text", "numeric", "numeric", "numeric", 
                                   "numeric", "numeric", "numeric", 
                                   "numeric", "numeric", "text", "text", 
                                   "text", "numeric", "numeric", "numeric"))
    ds_PeterCor <- merge(ds_PeterCor, ds_i, all = TRUE)
}

# Here information about group:
ds_PeterCor_Q <- read_excel("Data/Consistency_scores.xlsx")

names(ds_PeterCor_Q)[names(ds_PeterCor_Q) == "PPcode"] <- "ID"
names(ds_PeterCor)[names(ds_PeterCor) == "Code"] <- "ID"

ID_IN <- intersect(ds_PeterCor$ID, ds_PeterCor_Q$ID)
ds_PeterCor <- ds_PeterCor %>%
    filter(ID %in% ID_IN)
ds_PeterCor_Q <- ds_PeterCor_Q %>%
    filter(ID %in% ID_IN)

# Merge data with group and all dataset: 
ds_PeterCor <- ds_PeterCor_Q %>%
  left_join(ds_PeterCor, by = "ID")

# Rename to match:
names(ds_PeterCor)[names(ds_PeterCor) == "Group"] <- "group"
ds_PeterCor[ds_PeterCor$group == "SSS",]$group <- "Syn"
ds_PeterCor[ds_PeterCor$group == "control",]$group <- "Ctl"
names(ds_PeterCor)[names(ds_PeterCor) == "MouseClick.RESPCursorY"] <- "y"
names(ds_PeterCor)[names(ds_PeterCor) == "MouseClick.RESPCursorX"] <- "x"
names(ds_PeterCor)[names(ds_PeterCor) == "word"] <- "stimulus" # Will need to translate to english!
names(ds_PeterCor)[names(ds_PeterCor) == "BlockList.Cycle"] <- "repetition"

ds_PeterCor$width  <- 1920 # "screen with display resolution set to 1920 x 1080, controlled by a Dell computer running Windows 7."
ds_PeterCor$height <- 1080
ds_PeterCor$dataSource <- "PeterCor"

ds_PeterCor <- ds_PeterCor %>%
  select(all_of(ColNames_ds))

# Translate, weekdays:
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "maandag"]   <- "Monday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "dinsdag"]   <- "Tuesday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "woensdag"]  <- "Wednesday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "donderdag"] <- "Thursday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "vrijdag"]   <- "Friday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "zaterdag"]  <- "Saturday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "zondag"]    <- "Sunday"

# Translate, Monthc:
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "januari"]   <- "January"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "februari"]  <- "February"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "maart"]     <- "March"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "april"]     <- "April"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "mei"]       <- "May"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "juni"]      <- "June"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "juli"]      <- "July"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "augustus"]  <- "August"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "september"] <- "September"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "oktober"]   <- "October"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "november"]  <- "November"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "december"] <- "December"

# They added 2 numbers that are not in the other datasets:
ds_PeterCor <- ds_PeterCor %>%
    filter(stimulus != "50") %>%
    filter(stimulus != "100")

# MERGE  
ds   <- merge(ds,ds_PeterCor, all = TRUE)

# Questionnaire:
names(ds_PeterCor_Q)[names(ds_PeterCor_Q) == "Group"] <- "group"
ds_PeterCor_Q[ds_PeterCor_Q$group == "SSS",]$group <- "Syn"
ds_PeterCor_Q[ds_PeterCor_Q$group == "control",]$group <- "Ctl"

ds_PeterCor_Q$dataSource <- "PeterCor"

ds_PeterCor_Q <- ds_PeterCor_Q %>%
  select(ID, group, `Consistency All`, `Consistency Days`, `Consistency Months`,`Consistency Numbers`, dataSource)

# Append rows:
ds_Q <-  bind_rows(ds_Q,ds_PeterCor_Q)

# Tidy workspace:
rm(ds_PeterCor_Q,ds_PeterCor,ds_i, ID, fn,ColNames_ds,i,ID_IN)
```

```{r MergeWard2}
# This data was provided by Jamie by mail. The data is prepared in a separeate file
# Preserve "?" in the colnames: 
ds_Ward2 <- read.csv2("Data/WardData2.csv", fileEncoding = "UTF-8", check.names = FALSE)
ds_Ward2$ID <- ds_Ward2$session_id 
ds_Ward2$dataSource <- "Ward2"

# Add Cond
ds_Ward2$Cond <- NaN
ds_Ward2$Cond[ds_Ward2$stimulus %in% c("1","2","3","4","5","6","7","8","9","0")] <- "number"
ds_Ward2$Cond[ds_Ward2$stimulus %in% c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")] <- "weekday"
ds_Ward2$Cond[ds_Ward2$stimulus %in% c("January", "February", "March", "April", "May","June","July","August","September","October","November","December")] <- "month"

# Now I somehow need to define the groups
# If a response in the questionnaire then syn, if not then control.
ds_Ward2$group <- NaN
ds_Ward2$group[is.na(ds_Ward2$`Q1 . Some people routinely think about sequences as arranged in a particular spatial configuration (as in the examples below), do you think this might apply to you? (1=strongly agree, 5= strongly disagree)`)] <- "Ctl"
ds_Ward2$group[!is.na(ds_Ward2$`Q1 . Some people routinely think about sequences as arranged in a particular spatial configuration (as in the examples below), do you think this might apply to you? (1=strongly agree, 5= strongly disagree)`)] <- "Syn"

colNWard2 <- colnames(ds_Ward2)

ds_Ward2_Q <- ds_Ward2 %>%  
  select(c(ID, dataSource, colNWard2[colNWard2 %in% colnames(ds_Q)])) %>% 
  group_by(ID) %>%
  filter(row_number() == 1)

ds_Ward2_Q <- as.data.frame(ds_Ward2_Q)
ds_Ward2_Q$`questionnaire score` <- ds_Ward2_Q[,4] +  
  rowSums(ds_Ward2_Q[,21:24]) +
  rowSums(6 - ds_Ward2_Q[,25:28])

## Merge
ds         <- merge(ds,ds_Ward2, all = TRUE)
ds_Q       <- merge(ds_Q,ds_Ward2_Q,all = TRUE)

rm(ds_Ward2,ds_Ward2_Q)
```

```{r Wrangle data}
# Now we can enrich our dataset and process  several checks
# Add Condition, i.e. stim type:
ds$Cond <- NaN
ds$Cond[ds$stimulus %in% c("1","2","3","4","5","6","7","8","9","0")] <- "number"
ds$Cond[ds$stimulus %in% c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")] <- "weekday"
ds$Cond[ds$stimulus %in% c("January", "February", "March", "April", "May","June","July","August","September","October","November","December")] <- "month"

# Add stimulus repetition number
ds <- ds %>%
  group_by(ID, stimulus) %>%
  arrange(ID, stimulus, .by_group = TRUE) %>%
  mutate(repetition = row_number()) %>%
  ungroup()

# few trials that have more than 3 repetitions somehow:
ds <- ds %>% filter(repetition <= 3)

# Factorialize group variable to avoid confusions:
ds$group   <- as.factor(ds$group)
ds_Q$group <- as.factor(ds_Q$group)
```

```{r, FilterTrials}
# This is potentially a critical part for the results. Need to decide:
# (1) how to treat same coordinate responses, i.e. when ID clicks on the same position (see also Rothen and Ward)
# (2) how to treat missing data (i.e. NA & NaN)

# For (1), since z-score might fail (i.e. if all same coordinates) -->  turn same coordinates to NaN
# For (2), since sf does not accept missing data -->  we exclude those responses.
# This leads to some stimuli having only 2 repetitions (i.e. ID %in% 4_JB) AND some ID having to few pairs.
# Since polygons need at least 4 coordinate pairs. --> check enough data per repetition

# Important. To reproduce Rothen, the X = -1 and Y = -1 need to be interpreted as missing data (NaN):
ds$x[ds$x == -1] <- NaN
ds$y[ds$y == -1] <- NaN

# Detect ID has the same coordinates across Conditions:
ds <- ds %>%
    group_by(ID, Cond) %>% 
    mutate(SameCoord_x = length(unique(x[!is.na(x)])) == 1,SameCoord_y = length(unique(y[!is.na(y)])) == 1)

# NaNification:
ds$x[ds$SameCoord_x] <- NaN
ds$y[ds$SameCoord_y] <- NaN
ds$x[is.na(ds$x)]    <- NaN  # In VanPeters, some NA data all as NaN
ds$y[is.na(ds$y)]    <- NaN 

N1 <- length(unique(ds$ID)) # Track if ID loss

# Remove NaN from the dataset:
All_n <- length(ds$x) # Track trial loss
ds <- ds %>%
  filter(!is.na(x))
ds <- ds %>%
  filter(!is.na(y))
All_n2 <- length(ds$x)

N2 <- length(unique(ds$ID))  # Track if ID loss

# Check that each stimulus has 3 repetitions (3+2+1 = 6).
ds <- ds %>%
    ungroup() %>%
    group_by(ID, stimulus) %>%
    mutate(SumRep = sum(repetition)== 6)

ds <- ds[ds$SumRep,]
All_n3<- length(ds$x)

# Counts how many full conditions each ID has (should be 3):
tmp2 <- ds %>%
  ungroup() %>%
  group_by(ID, Cond, repetition) %>%
  filter(row_number() == 1) %>%
  summarize(ntrials = length(x)) %>%
  ungroup() %>%
  group_by(ID) %>%
  summarize(nCond = sum(ntrials)/3)

# Later on, we make polygons. These need at least 4 points. Hence we exclude ID's with less than 4 responses per condition:
tmp <- ds %>%
  ungroup() %>%
  group_by(ID, Cond, repetition) %>%
  summarize(ntrials = length(x))

ds <- ds %>%
  filter(!ID %in% tmp$ID[tmp$ntrials < 4])

N3 <- length(unique(ds$ID))

# Match ID to ds_Q:
IDlist <- unique(ds$ID)
ds_Q <- ds_Q %>%
  filter(ID %in% IDlist)
```

```{r Add_zs}
## Standardize/scale coordinates
ds <- ds %>%
  group_by(ID) %>% 
  mutate(x_zs = scale(x)) %>%
  mutate(y_zs = scale(y))
```

```{r ManualAdjust}

# Manually adjust pixels of invalid screen size
# Note: 29 since 29 stimuli
ds$height[ds$ID == 29324] <- 1080
ds$width[ds$ID == 32190 ] <- 1440

# This ID has unexpected changing screen settings, consider exclusion:
ds$width[ds$ID == 33168 ]  <- 308
ds$height[ds$ID == 33168 ] <- 149

ds$width[ds$ID == 35556 ]  <- 1439
ds$height[ds$ID == 35556 ] <- 734

ds$width[ds$ID == 48114 ]  <- 1593
ds$height[ds$ID == 48114 ] <- 671

ds$width[ds$ID == 59854]  <- 1366
ds$height[ds$ID == 59854] <- 663

ds$width[ds$ID == 63127]  <- 1920
ds$height[ds$ID == 63127] <- 880


ds$Screen_area <- ds$width*ds$height
```

# Methods

Code and data for “one-click” reproducibility of this manuscript
including the following analyses are available on git-hub
(<https://github.com/remLach/SpaceSequenceSynDiagnostic>). All the
analyses are conducted in R, using the following packages
`r pakret::pkrt_list("pROC", "sf")` and for visualization
`r pakret::pkrt_list("ggplot2", "ggVennDiagram")`.

## *Datasets and Participants*

To base the optimization of the consistency test on the largest possible
sample we looked for accessible datasets that used the same consistency
tasks on groups of synesthetes and controls. We found four datasets that
met these criteria. Two datasets were collected in laboratory settings
@rothen2016 (see also
[https://osf.io/6hq94](https://osf.io/6hq94/files/osfstorage)) and
@vanpetersen2020 (see also
<https://data.ru.nl/collections/di/dcc/DSC_2018.00019_653>). The two
others came from on-line testing @ward2022a (see
also<https://osf.io/nu5v4>) and an additional dataset was gently
provided by private communications with Professor Ward. To match the
other datasets, stimuli of the weekdays and months categories were
translated from Dutch to English in the dataset from [@vanpetersen2020].
For the number category we included only digits from 0 to 9 that where
presented across all datasets (i.e. excluding the stimuli "50" and "100"
for numbers in @vanpetersen2020).

From the merged datasets, we kept `r N3` from the total `r N1`
participants. First, we excluded `r round(((All_n - All_n2)/All_n),2)` %
empty trials (i.e. skipped responses) including trials flagged for
having the same x or y coordinates across conditions and repetitions,
causing the depletion n = `r N1-N2` participants[as in @rothen2016;
@ward2018]. n = `r N2-N3` participants were excluded for having less
than 4 coordinate points since polygons need at least four coordinate.
Note therefore, that `r sum(tmp2$nCond < 3)` participants of the final
sample did not have responses in all three conditions X repetition
cases. From the final sample of N = `r length(ds_Q$ID)`,
`r sum(ds_Q$group == "Syn")` were synaesthetes and
`r sum(ds_Q$group == "Ctl")` controls, see @tbl-mytable01 and
@tbl-mytable02.

| Synaesthetes       | N before | N after | Age  |
|--------------------|----------|---------|------|
| [@rothen2016]      | 33       | 32      | 23.1 |
| [@vanpetersen2020] | 23       | 22      | 23.2 |
| [@ward2022a]       | 252      | 252     | 37.2 |
| Ward 2             | 90       | 90      | NA   |
| Total:             | 398      | 396     |      |

: Synesthetes {#tbl-mytable01 apa-note="Synesthetes. Demographics were
not provided in the dataset Ward 2. Before: before participant
exclusion"}

| Controls           | N before | N after | Age  |
|--------------------|----------|---------|------|
| [@rothen2016]      | 37       | 37      | 28.2 |
| [@vanpetersen2020] | 21       | 21      | 21.6 |
| [@ward2022a]       | 215      | 213     | 19.9 |
| Ward 2             | 18       | 18      | NA   |
| Total              | 291      | 289     |      |

: Controls {#tbl-mytable02 apa-note="Controls. Demographics were not
provided in the dataset Ward 2. Before: before participant exclusion"}

Regarding the synaesthets profiles, detailed participant information
such as questionnaire responses were only available from the two
datasets from the data by Professor Ward (i.e. a majority of
`r sum(ds_Q$dataSource %in% c("Ward","Ward2"))` participants). We
described the self-reported profiles for the stimulus categories used in
the consistency test (i.e. number, weekdays and month), see
@fig-myplot01. From this Venn-diagram we see that 240 of the SSS report
having spatial-forms for Numbers, Days and Month; 42 only Days and
Month; 21 only Number and Day. Also, 127 controls report spatial-forms
for either Days, Numbers, Months or combinations.

```{r}
#| label: fig-myplot01
#| fig-cap: "Venn diagram of the types of self-reported SSS"
#| apa-note: "Only data from Pr. Ward is represented here. SSS = Sequence-space synaesthesia"

# Load library

na.rm <- function(x){x <- x[!is.na(x)]}
ds_Q_W <- ds_Q %>% filter(dataSource %in% c("Ward","Ward2"))

ID_all    <- na.rm(ds_Q_W$ID[ds_Q_W$dataSource %in% c("Ward","Ward2")])
ID_group  <- na.rm(ds_Q_W$ID[ds_Q_W$group =="Syn"])
ID_number <- na.rm(ds_Q_W$ID[ds_Q_W$`Q2_numbers (1=present, 0=absent)` == 1] )
ID_days   <- na.rm(ds_Q_W$ID[ds_Q_W$`Q2_days  (1=present, 0=absent)` == 1] )
ID_month  <- na.rm(ds_Q_W$ID[ds_Q_W$`Q2_months  (1=present, 0=absent)`== 1] )

x <- list(SSS = ID_group, Number = ID_number, Day = ID_days, Month =ID_month)
# An euler diagram would be best, but there are no good solutions in r for now
ggVennDiagram(x, label = "count", set_color = c("black","orange","orange","orange")) +
    scale_fill_gradient(low = "white", high = "firebrick")
# If you enjoy more complex visualization than mental health:
# library(nVennR)
# v <- plotVenn(x)
# showSVG(v, outFile = "venn.html", systemShow = TRUE)
```

```{r, cleanup3}
# Clean up :
rm(ID_number,ID_days,ID_month, tmp, x, colNWard2 ,N1,N2,N3)
```

## Procedure

For the consistency task, each stimuli were presented randomly and
sequentially centrally on the screen. Participant were instructed to
click on the screen position where they visualize them. In
@vanpetersen2020 and @rothen2016 the participants were allowed to skip
responses.

Stimulus included here were 7 weekdays (Monday to Sunday), 12 months
(January to December) and 9 numbers (0 to 9). For all the data from
Professor Ward the stimuli were presented in randomized order with the
constraint that no stimulus was repeated until all unique stimuli (N =
29) had been presented once. The median display resolution was
`r median(ds$width)` X `r median(ds$height)`, with a maximum of
`r max(ds$width)` X `r max(ds$height)` and a minimum of
`r min(ds$width)` X `r min(ds$height)`.

## Analyses

### Consistency tests on the *stimulus level*

We reproduced the tests extracted from consistency tasks found in the
literature such as area and perimeter between repetitions [@rothen2016;
@ward2022a; @vanpetersen2020]. Formally, the area (see [@eq-area]) and
perimeter (see [@eq-perim]) are calculated from the triangle (i.e. given
three repetitions) formed by the *x* and *y* coordinates of the three
repetitions of each stimuli (i.e. *(x1, y1), (x2, y2), (x3, y3)*).

$$
Area = \frac{|x1y2 +x2y3 + x3y1 – x1y3 – x2y1 – x3y2|}{2}
$$ {#eq-area}

$$
Perimeter = \sqrt{(x2 - x1)^2 + (y2 – y1)^2} + \sqrt{(x3 - x2)^2 + (y3 –y2)^2} + \sqrt{(x1 - x3)^2 + (y1 – y3)^2}
$$ {#eq-perim}

Each stimuli’s area was then averaged by participants. The area metric
is in % of the screen area to be able to compare with the consistency
across studies using different screen sizes. To account for screen size
and spread variability across the datasets and participants, area and
perimeter were also calculated on individually normalized x and y
coordinates (*z-score*). These methods compute consistency metrics at
the stimulus level, the rationale is that consistent responses would
show less variability between repetitions. For example, smaller area or
perimeter of the triangle formed indicate more variability hence more
consistent responses between the three repetitions.

```{=html}
<!--#
# Excluded this method because it is chronophage, does not work well (i.e. ROC). 

### Permuted consistency

Higher-order analytical methods such as permutation-based methods were
also used as metrics. We reproduced @root2021 permutation test for
colour-grapheme synaesthesia and adapted it to SSS as in @ward2022a. For
each individual, the x and y coordinates were randomly permuted and the
areas were calculated. After 1000 permutations per individual, a z-score
was calculated with the observed means compared to the permuted
distribution [@eq-zsPermRoot]. The distribution of the permuted results
form an individual chance level distribution, hence the z-score reflects
where the observed area lies on the chance level distribution.

$$
Zscore = \frac{(Observed) – (Mean Permuted)}{(SD Permuted)}
$$ {#eq-zsPermRoot}

The rationale is that synaesthetes consistencies should be closer (?).
 I think I would exclude this one. The permutation was originally used setting an arbitrary threshold -->
```

### Consistency tests on the *space-form level*

For the novel consistency tests, we extracted geometrical features at
the spatial-form level for each repetition separately. Taking the
stimuli as an ordered sequence we can consider them as a geometrical
segments (i.e. open geometrical form) and polygons (i.e. closed
geometrical form), similarly as originally described in @galton1880.

We harnessed a geospatial analysis, `r pakret::pkrt("sf")` to extract
polyons and geometry-based features from participant’s 2D (x,y)
coordinate responses. This package allows, for example, to build and
analyse segments or polygons and then extract multiple geometrical
descriptors or features. Informed by the ordinality of the stimulus, we
defined segments and polygon by conditions and repetitions. The
rationale here was to determine whether, when considering the
spatial-forms formed by stimuli as ordered coordinates (i.e. as segments
or polygon) they remain consistent *between* repetitions for each
individuals. We extracted the topological validity, topological
simplicity and a compactness score for each polygons (i.e. 3 conditions
X 3 repetitions: 9 polygons per participants.

#### Segment self-intersections

The first consistency test extracted from the segments is
self-intersection. We considered the spatial-forms as a segment
constituted by the lines connecting consecutive stimuli separately for
each category and repetitions (i.e. Monday (repetition 1) `→` Tuesday
(repetition 1) `→` Wednesday (repetition 1), ect) and then counted the
number of times each segment self-intersect. The rationale is that SSS
should have less chance to produce that self-intersect than control. The
number of self-intersections were computed separately for each
repetitions and conditions and averaged per participants.

#### Topological validity and simplicity

We assessed topological validity using geometric validation test.
Topological validity tests if a polygons is well-formed and valid
according to the Open Geospatial Consortium (OGC) Simple Features
Specification [@herring2010]. A polygon is considered topologically
valid if it satisfies the following criteria: (1) if polygon rings are
simple (i.e., they do not touch or self-intersect), (2) boundary rings
to not cross (3) boundary rings may only touch points tangentially (4)
rings that define holes are contained within the exterior ring (5) the
polygon rings must not splits the polygon [@postgis]. Topological
simplicity evaluates whether a geometric object has a simple structure
without self-intersections or self-tangencies. According to the OGC
Simple Features Specification [@herring2010], a polygon boundary is
considered simple if it does not pass through the same point more than
once [@postgis]. For polygons, simplicity requires that each ring does
not self-intersect or self-touch. Note that simplicity is a condition
for validity: a polygon can be simple but still invalid (e.g., if an
interior ring extends outside the exterior ring).

For each participant we tested for topological validity and simplicity
across individual 3 categories (weekdays, month and numbers) and 3
repetitions, hence 9 space-forms or polygons. The binary outcome (1 = is
valid, 0 = invalid) was then averaged across all 9 forms. Hence, it can
span from 1 (all 9 are valid/simple) to 0 (none of the 9 forms are
valid/simple). Thus, the topological validity and simplicity score
ranges from 0 (none of the nine polygons are valid/simple) to 1 (all
nine polygons are valid). For example, if a participant’s polygon for
weekdays were valid across all three repetitions, months were valid in
two of three repetitions, and numbers were invalid in all repetitions,
the topological validity score would be 5/9 ≈ 0.56.

#### Compactness score

We also implemented a measure of compactness for the polygons. We used
the score developed to quantify the gerrymandering of political district
by @polsby1991. Gerrymandering is a politically motivated strategy of
redefining district's boundary to give the advantage to a political
party. This has the consequences to reduce the compactness of those
districts as would be expected if they where defined from geomorphology.
The compactness is calculated as the ratio of the area and perimeter of
a polygon. The rationale is that more compact forms (i.e. higher
Polsby–Popper scores) would be more likely in SSS than controls.
Formally, the Polsby–Popper score is calculated from the polygon's area
and perimeter [@eq-PP]:

$$
score = \frac{4\pi \cdot Area}{Perimeter^2}
$$ {#eq-PP}

#### Permutation-based feature extraction

Finally, we implemented permutations of the repetition order to derive a
permuted measure of consistency. Specifically, instead of using
repetitions in chronological order, we shuffled the repetition order
between stimulus presentation (i.e. Monday (repetition 2) `→` Tuesday
(repetition 3) `→` Wednesday (repetition 1), ect) and extracted the
geometrical feature from x,y coordinates of same stimulus categories in
non-chronological order The form-based tests computed before were
relying on the chronologically ordered repetitions. For example, when a
stimulus such as Monday was repeated three times, the coordinates for
the first presentation of Monday were always paired with the coordinates
for the first presentation of Tuesday to construct segments or polygons.
However, if synesthetic forms are truly consistent, they should remain
stable independently of the chronological order of stimulus
presentation. Hence, we permute the repetitions within each condition.
Specifically, for each participant and each category, we randomized the
response repetition order. For example segments for weekdays were
constructed with the x,y coordinates where randomly permuted across
repetition order, for example Monday (repetition 1), Tuesday (repetition
3), Wednesday (repetition 2), ect. We predicted that this permuted-based
consistency would yield to better classification performance (higher AUC
values) compared to chronologically ordered features. The rationale is
that genuine synaesthetes would have a consistently valid topology of
spatial-forms repetitions regardless of presentation order, whereas
controls would show random variation. The permutation scores were
calculated for topological validity and Polsby–Popper score.

### Receiver Operator Characteristic analyses

Each of the previous method's performance in classifying SSS from
controls was compared with Receiver Operator Characteristics (ROC)
analyses. Classification performance was quantified using the area under
the curve (AUC), with higher AUC values indicating better discrimination
between SSS and control groups. The optimal cutoffs were calculated
using Youden’s index[@youden1950]. Discriminant Power (DP) was used as
an additional estimate of discrimination performance according to the
optimal cutoff (@eq-DP). DP around 1 being interpreted as inefficient
discrimination.

$$ 
DP = \frac{\sqrt{3}}{\pi} (\log(\frac{sensitivity}{1−sensitivity}) + \log(\frac{specificity}{(1−specificity)}))  
$$ {#eq-DP}

The statistical comparison of ROC curves between the methods was
operated testing the difference of AUC between pairs of curves using
bootstrapping method (1000 times) since it allows to compare methods
with different directions. Each bootstrap results in the subtraction of
the AUC between the two ROC. Then the each of the AUC difference's (D)
are divided by the standard deviation of those difference and compared
to a normal distribution (@eq-D).

$$
D = \frac{AUC_1-AUC_2}{SD}
$$ {#eq-D}

Therefore we statistically compared all the AUC of each methods as the
difference to the highest AUC method.

```{r, triangleArea_fun}
# Define area calculation function
triangle_area <- function(x, y) {
  if(length(x) != 3 | length(y) != 3) return(NA)
  area <- abs(
    x[1]*y[2] + x[2]*y[3] + x[3]*y[1] -
    x[1]*y[3] - x[2]*y[1] - x[3]*y[2]
  ) / 2
  return(area)
}
```

```{r, area_Consitency}
## Same with z-scores:
ds <- ds %>%  
  group_by(ID, stimulus) %>%
  mutate(rep_area = triangle_area(x, y)) %>%
  ungroup()

## Merge
tmp_perID2 <- ds %>% 
    ungroup() %>% 
    group_by(ID) %>% 
    summarize(Area_perc = sum(rep_area, na.rm = TRUE)/(sum(!is.na(rep_area)))*100, Screen_area = unique(Screen_area))  %>%
  select(ID, Area_perc,Screen_area)

tmp_perID2$Area_perc <- tmp_perID2$Area_perc/tmp_perID2$Screen_area
tmp_perID2 <- tmp_perID2 %>% select(ID, Area_perc)
ds_Q <- merge(ds_Q,tmp_perID2,by = "ID")

feature_direction <- c("moreCtl")
rm(tmp_perID2)
```

```{r, area_zsConsitency}
## Same with z-scores:
ds <- ds %>%  
  group_by(ID, stimulus) %>%
  mutate(rep_area_zs = triangle_area(x_zs, y_zs)) %>%
  ungroup()

## Merge
tmp_perID2 <- ds %>% ungroup() %>% 
  group_by(ID) %>% 
  summarize(Area_zs = mean(rep_area_zs, na.rm = TRUE))  %>%
  select(ID, Area_zs)

ds_Q <- merge(ds_Q,tmp_perID2,by = "ID")
rm(tmp_perID2)

feature_direction <- c(feature_direction,"moreCtl")
```

```{r, perimFun}
### Perimeter function: 
triangle_perimeter <- function(x, y) {
  if(length(x) != 3 | length(y) != 3) return(NA)
  # Side lengths
  a <- sqrt((x[2] - x[1])^2 + (y[2] - y[1])^2)
  b <- sqrt((x[3] - x[2])^2 + (y[3] - y[2])^2)
  c <- sqrt((x[1] - x[3])^2 + (y[1] - y[3])^2)
  perimeter <- a + b + c
  return(perimeter)
}
```

```{r, perim}
## Compute triangle perimeter by group:
ds <- ds %>%  
  group_by(ID, stimulus) %>%
  mutate(triangle_perim = triangle_perimeter(x, y)*100/Screen_area) %>%
  ungroup()

## Summarize By ID:
tmp_ID <- ds %>%
  ungroup() %>% group_by(ID,group) %>%
  summarize(Perimeter_perc = mean(triangle_perim, na.rm = TRUE)) %>%
  select(ID, Perimeter_perc)

## Merge
ds_Q <- merge(ds_Q,tmp_ID,by = "ID")
feature_direction <- c(feature_direction,"moreCtl")
```

```{r, perim_zs}
## Compute triangle perimeter by group:
ds <- ds %>%  
  group_by(ID, stimulus) %>%
  mutate(triangle_perim_zs = triangle_perimeter(x_zs, y_zs)) %>%
  ungroup()

## Summarize By ID:
tmp_ID <- ds %>%
  ungroup() %>% group_by(ID,group) %>%
  summarize(Perimeter_zs = mean(triangle_perim_zs, na.rm = TRUE)) %>%
  select(ID, Perimeter_zs)

## Merge
ds_Q <- merge(ds_Q,tmp_ID,by = "ID")
feature_direction <- c(feature_direction,"moreCtl")
```

```{r RootPerm}
#| eval: false
#| include: false

# Code retrieved from OSF (adapted here):
### Create a simulated distribution of consistency.  Note that each time this is run it will give a slightly different answer due to the randomisation

IDlist <- unique(ds$ID)

simulated_consistency <- data.frame() 
observed_consistency <- data.frame() 

n <- 1000  # Total number of iterations
bar_width <- 50
update_points <- round(seq(1, length(IDlist), length.out = 200))

for(ID_n in 1:length(IDlist)) {
  
  if (ID_n %in% update_points || length(IDlist) == n) {
    percent <- ID_n / length(IDlist)
    num_hashes <- round(percent * bar_width)  # progress bar
    bar <- paste0("[", 
                  paste(rep("#", num_hashes), collapse = ""), 
                  paste(rep("-", bar_width - num_hashes), collapse = ""), 
                  "]")
    cat(sprintf("\r%s %3d%%", bar, round(percent * 100)))
    flush.console()
  }  # end progress bar

  ds_ID <- ds %>%
    filter(ID %in% IDlist[ID_n]) %>%
    select(ID,x,y, stimulus, Screen_area)
  
  observed_consistency[ID_n,1] <- unique(ds_ID$ID)
  observed_consistency[ID_n,2] <- ds_Q %>% filter(ID %in% IDlist[ID_n]) %>% select(Area_zs) # unique(ds_ID$Consistency)
  
  for (N_shuffle in 1:1000) {
    
    ## shuffle the indicidual data
    ds_ID$stim_shuffled <- sample(ds_ID$stimulus)
    shuffled <- ds_ID
    
    Stim_list <- unique(shuffled$stim_shuffled)
    
    shuffled <- shuffled %>%
      group_by(stim_shuffled) %>%
      mutate(area = triangle_area(x, y))
    
    simulated_consistency[N_shuffle,1] = sum((na.rm(shuffled$area))/(length(unique(shuffled$stim_shuffled))))*100/min((shuffled$Screen_area)) # Few ID have changed screen size during session.
      }
  
  ## calculate the p-value and z-score of the observed consistency
  observed_consistency[ID_n,3] <- mean(simulated_consistency[,1])
  observed_consistency[ID_n,4] <- sd(simulated_consistency[,1])
  observed_consistency[ID_n,5] <- (observed_consistency[ID_n,2] - observed_consistency[ID_n,3]) / (observed_consistency[ID_n,4])
  
  # Circumvent issue that some ID have area of 0. Since can't divide by 0 for the z-score.
  if(observed_consistency[ID_n,3] == 0){
    observed_consistency[ID_n,5] <- 0
  }
  
  if(!is.finite(observed_consistency[ID_n,5])){
      break()
      warning("z-score leads to non numeric results")
    }
}

colnames(observed_consistency) <- c('ID', 'Area_zs', 'mean_perm', 'SD_perm', 'z-score')

write.csv2(observed_consistency, "permuted_area_Root.csv")
```

```{r RootPermAdd}
#| eval: false
#| include: false
ds_perm_cons <- read.csv2("permuted_area_Root.csv")
ds_perm_cons$perm_zs <- ds_perm_cons$z.score

ds_Q <- right_join(ds_Q, 
                   ds_perm_cons %>% select(ID, perm_zs),
                   by = "ID")
feature_direction <- c(feature_direction,rep("moreCtl",1))

rm(ds_perm_cons, tmp_ID,tmp2)
```

```{r, selfInter_fun}

# Define function:
count_self_intersections <- function(x, y, verbose = TRUE) {
  n <- length(x)
  if (n < 4) {
    if (verbose) cat("Need at least 4 points to check for self-intersection.\n")
    return(0)
  }

  # Orientation function
  orientation <- function(p, q, r) {
    val <- (q[2] - p[2]) * (r[1] - q[1]) - (q[1] - p[1]) * (r[2] - q[2])
    if (is.na(val)) return(NA)
    if (val == 0) return(0)
    if (val > 0) return(1) else return(2)
  }

  # Check if q lies on segment pr
  on_segment <- function(p, q, r) {
    if (any(is.na(c(p, q, r)))) return(FALSE)
    q[1] <= max(p[1], r[1]) && q[1] >= min(p[1], r[1]) &&
      q[2] <= max(p[2], r[2]) && q[2] >= min(p[2], r[2])
  }

  # Main intersection check
  segments_intersect <- function(p1, p2, p3, p4) {
    o1 <- orientation(p1, p2, p3)
    o2 <- orientation(p1, p2, p4)
    o3 <- orientation(p3, p4, p1)
    o4 <- orientation(p3, p4, p2)

    if (any(is.na(c(o1, o2, o3, o4)))) return(FALSE)

    # General case
    if (o1 != o2 && o3 != o4) return(TRUE)

    # Special colinear cases
    if (o1 == 0 && on_segment(p1, p3, p2)) return(TRUE)
    if (o2 == 0 && on_segment(p1, p4, p2)) return(TRUE)
    if (o3 == 0 && on_segment(p3, p1, p4)) return(TRUE)
    if (o4 == 0 && on_segment(p3, p2, p4)) return(TRUE)

    return(FALSE)
  }

  count <- 0
  for (i in 1:(n - 2)) {
    for (j in (i + 2):(n - 1)) {
      if (j == i + 1) next  # skip adjacent segments

      p1 <- c(x[i], y[i])
      p2 <- c(x[i + 1], y[i + 1])
      p3 <- c(x[j], y[j])
      p4 <- c(x[j + 1], y[j + 1])

      if (segments_intersect(p1, p2, p3, p4)) {
        count <- count + 1
        if (verbose) {
          cat(sprintf("Intersection #%d: segments (%d-%d) and (%d-%d)\n", count, i, i+1, j, j+1))
        }
      }
    }
  }

  if (verbose) cat("Total crossings:", count, "\n")
  return(count)
}

```

```{r, Self_Intersections}
# Number of intersections for each ID X Cond X repetition
# Important! The dataset must be correctly informed about stimulus order.
ds <- ds %>% 
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ungroup() %>%
  group_by(ID, Cond,repetition) %>%
  mutate(nLineCross = (count_self_intersections(x,y, verbose = FALSE))) %>%
  arrange(ID) # ensure ds remains ordered by ID
```

```{r, selfIntertodsQ}
## Average self-intersections per ID by Cond X repetition
tmp <- ds %>%
    group_by(ID,Cond,repetition) %>%
    filter(row_number() == 1) %>% # keep only 1 row per IDXCondXrep
    ungroup() %>% group_by(ID) %>% # Average per ID
    summarise(SelfInter = mean(nLineCross)) %>%
    select(ID, SelfInter) 

## Merge
ds <- merge(ds, tmp,by = "ID")

ds_Q <- right_join(ds_Q, 
           (ds %>% 
                ungroup() %>% 
                group_by(ID) %>% 
                filter(row_number() == 1) %>%
                select(ID, SelfInter)), by = "ID")


feature_direction <- c(feature_direction,"moreCtl")
```

```{r, ds_segm}

# Turn off spherical geometry:
sf::sf_use_s2(FALSE)

# Explicitly enforce item order (i.e. ordinality):
ds_segm <- ds %>%
  group_by(ID, stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ungroup() %>%
  arrange(ID) %>%
  filter(!is.nan(x_zs), !is.nan(y_zs)) %>% # sf hates NaN! Not needed anymore,
  mutate(
    group = as.character(group),
    ID     = as.character(ID),
    Cond      = as.character(Cond),
    repetition     = as.integer(repetition),
    dataSource = as.character(dataSource)
  ) %>%
  group_by(ID, Cond, repetition,group) %>%
  summarise(
    geometry = st_sfc(st_linestring(as.matrix(cbind(x_zs, y_zs)))), # preserves order
    .groups = "drop"
  ) %>%
  st_as_sf(crs = NA)
```

```{r, poly area}
## Pass from segment to polygon
ds_poly          <- st_cast(ds_segm, "POLYGON")
ds_poly$area     <- st_area(ds_poly)

## Poly area
ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(areaPoly_GA = mean(area, na.rm = TRUE))

## Merge
ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, areaPoly_GA),
                   by = "ID")

feature_direction <- c(feature_direction,"moreSyn")
```

```{r, polyPeri}
ds_poly$perimeter    <- st_perimeter(ds_poly)

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(perimPoly = mean(perimeter, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, perimPoly),
                   by = "ID")

feature_direction <- c(feature_direction,"moreCtl")
```

```{r, polyisSimple}
# Might depend on cast:
# ds_poly          <- st_cast(ds_segm, "POLYGON", group_or_split = TRUE)

ds_poly$isSimple <- st_is_simple(ds_poly)

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(isSimple_GA = mean(isSimple, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, isSimple_GA),
                   by = "ID")

feature_direction <- c(feature_direction,"moreSyn")
```

```{r, topoValid}
ds_poly$isValidStruct <- st_is_valid(ds_poly, geos_method = "valid_structure")

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(isValidStruct_M = mean(isValidStruct))

ds_Q <- right_join(ds_Q,
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, isValidStruct_M),
                   by = "ID")

feature_direction <- c(feature_direction,"moreSyn")
```

```{r permuteAcrossRep}
#| eval: false
#| include: false

# All possible permutations per participant would be:
# 3 sets: weekdays (7) * numerals (10) * months (12)= 840
# 3 repetitions: 3! = 6
# 850*6 = 5040 possibilities. Since 100 permutation take about 30 minutes it would take a long time. Alredy with 1000 permutation should take 3 hours -.-'.

# Takes about 40 minutes on my machine. So pre-saved the output. 
set.seed(42)

ds <- ds %>%
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>%
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ungroup() %>%
  group_by(ID, group, Cond, repetition) %>%
  mutate(StimOrder = 1:length(stimulus))

# First define output matrix:
n_perms <- 500
perm_names <- paste0("n_perm", sprintf("%03d", 1:n_perms))  # X001 … X100

ds_perm <-  as.data.frame(
  matrix(NA_real_, nrow = length(unique(ds$ID)), ncol = n_perms,
         dimnames = list(unique(ds$ID), perm_names))
)

ds_perm_pp <-  as.data.frame(
  matrix(NA_real_, nrow = length(unique(ds$ID)), ncol = n_perms,
         dimnames = list(unique(ds$ID), perm_names))
)

ID_list <- unique(ds$ID)
total = length(ID_list)*n_perms
pb <- txtProgressBar(min = 0, max = total, style = 3)
k <- 0


for(Perm_n in 1:n_perms){
  perm_here <- perm_names[Perm_n]
  for(ID_n in 1:length(unique(ds$ID))){
    
    ################ Extract data per ID: ################
    ds_ID <- ds %>%
      filter(ID %in% ID_list[ID_n]) %>%
      select(ID, group, stimulus, Cond, nLineCross, repetition, StimOrder,x_zs,y_zs,SelfInter)
    
     ################ Apply permutation (sample) across repetitions################
     ds_ID <- ds_ID %>%
      group_by(stimulus) %>%
      mutate(repetition_perm = sample(repetition)) 
    
    ################ Pass to sf ################
    ds_ID_segm <- ds_ID %>%
      group_by(ID, stimulus) %>%
      arrange(stimulus) %>%
      arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
      arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
      filter(!is.nan(x_zs), !is.nan(y_zs)) %>% # sf hates NaN! Not needed anymore, NaN's are managed above in the code.
      ungroup() %>%
      arrange(ID) %>%
      mutate(
        group = as.character(group),
        ID     = as.character(ID),
        Cond      = as.character(Cond),
        repetition_perm     = as.integer(repetition_perm)
      ) %>%
      group_by(ID, Cond, repetition_perm,group) %>%
      summarise(
        geometry = st_sfc(st_linestring(as.matrix(cbind(x_zs, y_zs)))), # preserves order
        .groups = "drop"
      ) %>%
      st_as_sf(crs = NA) 
    
    ################ Convert to poly and compute validity ################
    ds_ID_poly               <- st_cast(ds_ID_segm, "POLYGON")
    ds_ID_poly$isValidStructPerm <- st_is_valid(ds_ID_poly, geos_method = "valid_structure")
    
    ds_ID_poly <- ds_ID_poly  %>%
      mutate(isValidStructPerm = mean(isValidStructPerm, na.rm = TRUE)) %>%
      filter(row_number() == 1)
    
    ds_ID_poly <- ds_ID_poly %>%
      mutate(
        area = st_area(geometry),
        perim = st_length(st_boundary(geometry)),
        pp = as.numeric((4 * pi * area) / (perim^2))
      )  %>%
      filter(row_number() == 1)
    
    ################ save the perumted results ################
        if(!rownames(ds_perm[ID_n,]) == ID_list[ID_n]){
      warning("ID names do not match")
    }
    
    ds_perm[ID_n,Perm_n] <- ds_ID_poly$isValidStructPerm
    ds_perm_pp[ID_n,Perm_n] <- ds_ID_poly$pp
    k <- k + 1
    setTxtProgressBar(pb, k)
  }
}

# Now average the permutations: 
ds_perm$ID <-rownames(ds_perm)
ds_perm$group <- ds_perm$ID  %in% ds_Q$ID[ds_Q$group == "Ctl"] # is true if Ctl
ds_perm$group[ds_perm$group]  <- "Ctl"
ds_perm$group[ds_perm$group == "FALSE"] <- "Syn"

ds_perm_pp$ID <-rownames(ds_perm_pp)
ds_perm_pp$group <- ds_perm_pp$ID  %in% ds_Q$ID[ds_Q$group == "Ctl"] # is true if Ctl
ds_perm_pp$group[ds_perm_pp$group]  <- "Ctl"
ds_perm_pp$group[ds_perm_pp$group == "FALSE"] <- "Syn"

write.csv2(ds_perm, "permuted_isValid.csv")
write.csv2(ds_perm_pp, "permuted_pp.csv")
```

```{r, AveragePermute}
ds_perm <- read.csv2("Data/permuted_isValid.csv")

## ATTENTION; RECOMPUTE THE PERMUTATION WITH THE ACTUAL DATASET then remove this
ds_perm <- ds_perm %>% filter(ID %in% ds_Q$ID)
ds_Q   <- ds_Q %>% filter(ID %in% ds_perm$ID)
# read.csv2("permuted_isValid.csv")
ds_Perm_perID <-  ds_perm %>% 
  rowwise() %>% 
  mutate(isValid_perm_M = mean(c_across(n_perm001:n_perm100)), isValid_perm_Med = median(c_across(n_perm001:n_perm100))) %>% 
  select (ID, group, isValid_perm_M, isValid_perm_Med) # removed isValid_Med_perm_ID since median always underperformed the average


ds_Q <- right_join(ds_Q, 
                   ds_Perm_perID %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% arrange(ID) %>% filter(row_number() == 1) %>% select(ID, isValid_perm_M),
                   by = "ID")
feature_direction <- c(feature_direction,rep("moreSyn",1))
```

```{r, pp_test}
## Polsby–Popper test / Iso-Perimetric Quotient
ds_poly <- ds_poly %>%
    mutate(
    area = st_area(geometry),
    perim = st_length(st_boundary(geometry)),
    pp = as.numeric((4 * pi * area) / (perim^2))
  ) 

ds_Q <- right_join(ds_Q,
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% summarize(pp_M = mean(pp, na.rm = TRUE)) %>% select(ID, pp_M),
                   by = "ID")

feature_direction <- c(feature_direction,"moreSyn")
```

```{r, AveragePermutepp}
ds_perm_pp <- read.csv2("Data/permuted_pp.csv")

ds_Perm_perID <-  ds_perm_pp %>% 
  rowwise() %>% 
  mutate(pp_perm_M = mean(c_across(n_perm001:n_perm100), na.rm = TRUE), pp_perm_Med = median(c_across(n_perm001:n_perm100))) %>% 
  select (ID, group, pp_perm_M, pp_perm_Med) # removed isValid_Med_perm_ID since median always underperformed the average


ds_Q <- right_join(ds_Q, 
                   ds_Perm_perID %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% arrange(ID) %>% filter(row_number() == 1) %>% select(ID, pp_perm_M),
                   by = "ID")
feature_direction <- c(feature_direction,rep("moreSyn",1))
```

```{r GLM}
#| eval: false
#| include: false

## Remov the GLM because it is tricky to compare with other features. Also the difference betwen datasets seems to be a biggest problem now. I also don't know how I would discuss it.

# Here was the text:
# Finally, we used General Linear Model (GLM) on the two features with the best AUC and add the prediction as an additional feature. A GLM could provide a formula where multiple features could be combined in order to optimize AUC.

Model <- glm(group ~  Area_zs + isValid_perm_M, data = ds_Q, family = binomial)
pred = predict(Model)
ds_Q$GLM_Valid_areazs <- pred

feature_direction <- c(feature_direction,rep("moreSyn",1))

Model1 <- glm(group ~  perimPoly + pp_perm_M, data = ds_Q, family = binomial)
pred1 = predict(Model1)
ds_Q$GLM_perimPoly_ppPerm <- pred1

feature_direction <- c(feature_direction,rep("moreSyn",1))
```

```{r, cleanupVar}
rm(ds_perm, ds_Perm_perID, tmp, All_n, All_n2, All_n3, ID_all, ID_group)
```

```{r}
#| label: fig-myplot00
#| fig-cap: "Design for novel method for consistency test: from between trial varibility to Segment and polygons formed by the space-forms using geography tools."
#| apa-note: "Example participant for months"
#| fig-format: png
#| dpi: 300
#| fig-pos: "H"

gp1 <- ds %>%
  filter(ID %in%  "1025_MaJo") %>%
  filter(Cond %in% "month")  %>%
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>% # Start ggplot
  ggplot(aes(x = x_zs, y = y_zs, group = stimulus, label = stimulus, fill = stimulus)) +
  geom_polygon(alpha = 0.4) +
  geom_point(size = 0.05, alpha = 0.5) +
  geom_text(aes(x = x_zs, y = y_zs), size = 1.5, alpha = 0.3) +
  scale_fill_discrete(guide="none") +
  coord_equal() +
  theme_void() + 
  labs(title = "Stimulus level") +
  theme(plot.title = element_text(hjust = 0.5)) 

gp2 <- ggplot() +
  annotate("segment",
           x = 0, xend = 0.5, y = 0, yend = 0,
           linewidth = 1,
           arrow = grid::arrow(type = "closed")) +
    theme_void()

gp3 <- ds_segm %>%
  filter(ID %in%  "1025_MaJo") %>%
  filter(Cond %in% "month")  %>%
  ggplot(aes(colour = as.factor(repetition))) +
  geom_sf(alpha = 0.4) +
  guides(colour="none") +
  theme_void() + 
  labs(title = "Space-form Segment") +
  theme(plot.title = element_text(hjust = 0.5)) 

gp4 <- ggplot() +
  annotate("segment",
           x = 0, xend = 0.5, y = 0, yend = 0,
           linewidth = 1,
           arrow = grid::arrow(type = "closed")) +
    theme_void()

gp5 <- ds_poly %>%
  filter(ID %in%  "1025_MaJo") %>%
  filter(Cond %in% "month")  %>%
  ggplot(aes(fill = as.factor(repetition))) +
  geom_sf(alpha = 0.4) +
  guides(fill="none") +
  theme_void() + 
  labs(title = "Polgyon") +
  theme(plot.title = element_text(hjust = 0.5)) 

plot_grid(gp1,gp2,gp3,gp4,gp5, 
          nrow = 1)

```

# Results {#sec-results}

```{r, defineDPfun}
DP <- function(sensitivity, specificity){
  sensitivity = sensitivity/100
  specificity = specificity/100
  return(sqrt(3/pi)*(log(sensitivity/(1-sensitivity)) + log(specificity/(1-specificity))))
}
```

```{r, InitializeROC_fun}
## Define function to compute ROC:
Comp_ROC <- function(data, group_col, feature, ID, Ord){
  # Ord: must be: moreSyn or moreCtl. Wheter feature value is moer in Syn or ctl.
  if(!setequal(levels(data[[group_col]]), c("Syn","Ctl"))){
    return(warning("group name must inculde: Syn Ctl"))
    break
  }
  
  if (Ord == "moreSyn") {
    Dirhere <- "<"
  } else if (Ord == "moreCtl") {
    Dirhere <- ">"
  } else {
    return(warning("Ord must be moreSyn or moreCtl"))  
  }
  
  ################ ROC analyses ################ 
  ROC_here <- pROC::roc(data[[group_col]] ~ data[[feature]], data, 
                  direction= Dirhere,
                  percent=TRUE,
                  ci=TRUE, boot.n=100, ci.alpha=0.9, stratified=FALSE,
                  )
  
  # Best threshold using Youden's J
  best_coords <- pROC::coords(ROC_here, "best", 
                              ret = c("accuracy","threshold", "sensitivity","specificity","ppv","npv"), 
                              best.method = "youden")
  

  auc_val <- as.numeric(pROC::auc(ROC_here))
  ci_auc  <- ci.auc(ROC_here)
  ciobj   <- pROC::ci.se(ROC_here, # CI of sensitivity
               specificities=seq(0, 100, 5)) # over a select set of specificities

  new_row <- data.frame(
    Feature   = feature,
    AUC        = round(auc_val, 4),
    DP         = DP(as.numeric(best_coords[["sensitivity"]]),as.numeric(best_coords[["specificity"]])),
    threshold  = as.numeric(best_coords[["threshold"]]),
    sensitivity= as.numeric(best_coords[["sensitivity"]]),
    specificity= as.numeric(best_coords[["specificity"]]),
    # ci_low     = as.numeric(ci_auc[1]),
    # ci_high    = as.numeric(ci_auc[3]),
    stringsAsFactors = FALSE
  )
  
  ################ Contingency table ################ 
    if (Ord == "moreSyn") {
    data$diagnosis <- ifelse(data[[feature]] >= best_coords$threshold,  "Test : Syn", "Test : Ctl")
  } else if (Ord == "moreCtl") {
    data$diagnosis <- ifelse(data[[feature]] >= best_coords$threshold,  "Test : Ctl","Test : Syn")
  } else {
    warning("Ord must be moreSyn or moreCtl")  
  }
  
  tab_counts <- table(data[[group_col]], data$diagnosis)
  
  tab_percent <- prop.table(tab_counts, margin = 1) * 100
  
  result <- matrix(
    paste0(tab_counts, " (", round(tab_percent, 1), "%)"),
    nrow = nrow(tab_counts),
    dimnames = dimnames(tab_counts)
  )
  
  ################ General description ################ 
  # Attention, na removed  
  Descr_table <- data %>%
    group_by(!!sym(group_col)) %>%
    summarize(n = length(unique(!!sym(ID))), Mean = mean(!!sym(feature), na.rm = TRUE), SD = sd(!!sym(feature), na.rm = TRUE))

  ################ Return outputs ################ 
  return(list(ROC_properties = new_row, Coningency_table =result, Descr_table = Descr_table, ROC_curve = ROC_here, CI = ciobj))
}
```

```{r compROCallFeat}
#| message: false
#| warning: false

all_cols <- colnames(ds_Q)
feature_list      <- all_cols[42:length(all_cols)]

if(length(feature_list) != length(feature_direction)){
  warning("Attention unmatch feature list and direction")
  message("Attention unmatch feature list and direction")
}
Feats <- as.matrix(t(rbind(feature_list,feature_direction)))

# Now we will loop through features: INITIALIZE ROC_curves
ROC_curves <- list()
ROC_contingency <- list()

# Initialize dataframe to collect relevant ROC values:
All_ROC <- data.frame(Feature=character(),
                      AUC         = character(),
                      threshold   = integer(),
                      sensitivity = integer(),
                      specificity = integer(),
                      ppv         = integer(),
                      npv         = integer(),
                      high_ci     = integer(),
                      low_ci      = integer(),
                      # power       = integer(),
                      stringsAsFactors=FALSE)

# Loop by features:
for(fit_n in 1:length(Feats[,1])){
  ROC_here <- Comp_ROC(ds_Q, "group", Feats[[fit_n,1]],"ID", Feats[[fit_n,2]])
  
  ROC_curves      <- c(ROC_curves,list(ROC_here$ROC_curve,ROC_here$ROC_properties$Feature))
  All_ROC         <- rbind(All_ROC, ROC_here$ROC_properties)
  ROC_contingency <- c(ROC_contingency,list(ROC_here$Coningency_table,ROC_here$ROC_properties$Feature))
}
```

```{r, AUCtest}
# Multiple comparison issue (need ot adjust p val):
# AUC_test  <- roc.test(ds_Q$group, ds_Q$isValid_perm_M, ds_Q$Perimeter_zs,  boot.n=1000, method = "bootstrap")
# AUC_test2 <- roc.test(ds_Q$group, ds_Q$Perimeter_zs, ds_Q$isValidStruct_M,  boot.n=1000, method = "bootstrap")
# AUC_test3 <- roc.test(ds_Q$group, ds_Q$isValid_perm_M, ds_Q$isValidStruct_M,  boot.n=1000, method = "bootstrap")

# Compare ALL the features and adjust, but leads to the same results.
nFeat    <- length(ROC_curves)
Odd_idx  <- seq(1,nFeat,by=2)
Pair_idx <- seq(2,nFeat,by=2)

ROC_listed <- list()
Feattype <- c()
for(i in 1:nFeat){
  ROC_listed[[i]] <- ROC_curves[[Odd_idx[i]]]
  Feattype <- c(Feattype,ROC_curves[[Pair_idx[i]]])
}

allAUC  <- sapply(ROC_listed, auc)
sel_auc <- which(allAUC == sort(allAUC, decreasing = TRUE)[1])
toCompare_roc <- ROC_listed[[sel_auc]] 

statistic <- sapply(seq_along(ROC_listed), function(i) {
  if (i == sel_auc) return(NA)
  roc.test(toCompare_roc, ROC_listed[[i]],  boot.n=1000,method = "bootstrap")$statistic
})

pvals <- sapply(seq_along(ROC_listed), function(i) {
  if (i == sel_auc) return(NA)
  roc.test(toCompare_roc, ROC_listed[[i]],  boot.n=1000,method = "bootstrap")$p.value
})

# Adjust for multiple comparison: the order is Feattype
pvals_adj <- p.adjust(pvals, method = "fdr") # or fdr since exploratory, lead to the same.

Results_ROCcomp <- data.frame(
  Feature = Feattype,
  D = statistic, 
  pvals = pvals_adj,
  isSign = (pvals_adj <.05)
)
```

```{r}
#| label: fig-myplot2
#| fig-cap: "Receiver Operating Characteristic (ROC) curves of all tests"
#| apa-note: "Grey line indicates chance level"

nFeat <- length(ROC_curves)
Odd_idx  <- seq(1,nFeat,by=2)
Pair_idx <- seq(2,nFeat,by=2)

ROC_listed <- list()
Feattype <- c()
for(i in 1:nFeat){
  ROC_listed[[i]] <- ROC_curves[[Odd_idx[i]]]
  Feattype <- c(Feattype,ROC_curves[[Pair_idx[i]]])
}

Feattype <- recode(Feattype, 
                   "Area_perc" = "Area [screen size %]",
                   "Area_zs" =  "Area [z-score]", 
                   "Perimeter_perc" = "Perimeter [screen size %]",
                   "Perimeter_zs" = "Perimeter [z-score]",
                   "SelfInter" = "Segment self-intersections [n/9]",
                   "areaPoly_GA" = "Area of the polygon [z-score]",
                   "perimPoly" = "Perimeter of the polygon [z-score]" ,   
                   "isSimple_GA" = "Simple topology [boolean/9]",
                   "isValidStruct_M" = "Valid structure [boolean/9]",
                   "isValid_perm_M" = "Permuted valid structure [boolean/9]",
                   "pp_M" = "Polsby–Popper [score/9]",
                   "pp_perm_M" = "Permuted P-P [score/9]"
                  )


ggroc(ROC_listed) +
  geom_segment(aes(x = 0, xend = 100, y = 100, yend = 0),
               color="grey", size = 0.01) +
  scale_color_manual(labels = Feattype,values =pals::cols25(nFeat/2)) + # pals::brewer.pastel1(nFeat)
  papaja::theme_apa()+
  theme(legend.title = element_blank()) +
  coord_equal() +
  theme(legend.position="bottom",legend.text=ggplot2::element_text(size=7))
```

```{r}
# For tables, use gt() you will save the time I wasted. 
# Avoid gt() and kableExtra(), my issue was that the caption was in the middle of the first line.
# I tried all papaja::apa_table, lost much time (cross-referencing becomes a nightmare)
# Apa quarto is lacking a way to handle column spanning 

All_ROC_round <- All_ROC %>% 
    mutate_if(is.numeric, round,2)
All_ROC_round <- All_ROC_round[order(All_ROC_round$AUC, decreasing = TRUE), ]

All_ROC_round <- All_ROC_round %>% 
  mutate(Feature = recode(Feature, 
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" =  "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,   
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" = "Permuted valid structure [boolean/9]",
                          "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]"
  ))

SortFeat <- All_ROC_round$Feature # Pass further the AUC sorted features
```

```{r}
#| label: tbl-mytb03 
#| tbl-cap: "ROC analyses of all tests"
#| ft.align: left

# library(flextable)
# All_ROC_round |>
# flextable(All_ROC_round) %>%
#   flextable::theme_apa() |>
#   autofit()
   # kableExtra::kbl(booktabs = TRUE) 
   # gt()
# "AUC = Area Under the Curve. DP = Discrimination Power."
# papaja::apa_table(All_ROC_round, 
#                  caption = NULL,
#                   label = NA)
knitr::kable(All_ROC_round)
```

```{r}
#| label: tbl-mytb04
#| tbl-cap: "Descriptives of each test"
#| ft.align: left
#| apa-notes: "Mean (Standart Deviation) of all the consistency tests"

# Well this should table is present higher in the text, however it needs to be informed by the ROC analyisis (because I want the feature order to always be sorted by AUC).
feat_cols <- colnames(ds_Q[,42:length(colnames(ds_Q))])

fmt_ms <- function(x, digits = 2) {
  m    <- mean(x, na.rm = TRUE)
  Med  <- median(x, na.rm = TRUE)
  sd   <- stats::sd(x, na.rm = TRUE)
  paste0(
         formatC(m,  format = "f", digits = digits),
         " (",
         formatC(sd, format = "f", digits = digits),
         ")"
  )
}

tmp <- ds_Q %>%
    select(c(group, feat_cols)) %>%
    pivot_longer(cols = all_of(feat_cols),
                 names_to = "Feature",
                 values_to = "Value") %>%
    group_by(group,  Feature) %>%
    summarise(`M *(SD)*` = fmt_ms(Value), .groups = "drop") %>%
    pivot_wider(names_from = c(group),
                values_from = c(`M *(SD)*`)) 

tmp <- tmp %>%mutate(Feature = recode(Feature, 
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" =  "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,   
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" = "Permuted valid structure [boolean/9]",
                           "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]"
  ))

Descr_table <- tmp[match(All_ROC_round$Feature, tmp$Feature), ]

Descr_table_zs <- ds_Q %>%
  select(feature_list, group,ID) %>%
  mutate_at(feature_list,scale) %>%
  pivot_longer(cols = all_of(feat_cols),
               names_to = "Feature",
               values_to = "Value") %>%
  group_by(group,  Feature) %>%
  summarise(`M *(SD)*` = fmt_ms(Value), .groups = "drop") %>%
  pivot_wider(names_from = c(group),
              values_from = c(`M *(SD)*`)) 


Descr_table_zs <- Descr_table_zs %>% mutate(Feature = recode(Feature, 
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" =  "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,   
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" = "Permuted valid structure [boolean/9]",
                           "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]"
  ))

Descr_table_zs <- Descr_table_zs[match(All_ROC_round$Feature, Descr_table_zs$Feature),] %>%
  rename("Control [zs]" = "Ctl", "SSS [zs]" = "Syn")

# knitr::kable(merge(Descr_table, Descr_table_zs, by = "Feature",sort = FALSE),
#     col.names = c("Feature", "Controls", "Sequence-Space Synaesthetes", "Controls Median[zs]", "Sequence-Space Synaesthetes Median[zs]"), align ="c") 

tbl_out <- merge(Descr_table, Descr_table_zs, by = "Feature",sort = FALSE) %>% 
  rename("Control" = "Ctl", "SSS" = "Syn")

# I will not lie, it was painful to build this table:
# Correction: increasing pain did not improve the table. Got a latex error.  
# flextable(tbl_out) %>%
#   separate_header() |>
#   border_remove() |>
#   add_header_row(values = c("","Raw", "Standardized [zs]"), colwidths = c(1,2, 2)) %>%
#   hline(i = 1, part = "header",j = 2:5) %>%
#   flextable::theme_apa() |>
#   autofit() 
knitr::kable(tbl_out)

rm(tmp)
# Sanity Check: The descriptives for Area_perc calculated here are the ~same as in Ward
## Here:
# ds_Q %>% group_by(group) %>% filter(dataSource %in% "Ward") %>% summarize(M = mean(Area_perc, na.rm = TRUE), sd = sd(Area_perc, na.rm = TRUE))
## Ward
# ds_Q %>% group_by(group) %>% filter(dataSource %in% "Ward") %>% summarize(M = mean(consistency_score, na.rm = TRUE), sd = sd(consistency_score, na.rm = TRUE))
```

Descriptively, the triangle area (in % screen size) between repetition
for SSS was surprisingly larger than in previous studies: 0.26% (*SD =
0.52%*) compared to 0.15% (*SD = 0.13 %*) in @rothen2016 and 0.14 % (*SD
= 0.17*) % in @ward2018, see @tbl-mytb04. These values also differ
between the available datasets (See @fig-apx08). However, those
differences could be mitigated by using individually standardized
coordinates, see @tbl-mytb04.

## ROC analyses

All ROC were compared to the permuted validity method, since it led the
highest AUC (`r round(max(All_ROC$AUC),2)`), all p-values false
discovery rate (FDR) adjusted. ROC analyses are summarized in
@tbl-mytb03 and @fig-myplot2. Interestingly, while the permuted validity
leads to significantly larger AUC than the Polsby-Popper method, (*D* =
`r round(Results_ROCcomp[Results_ROCcomp$Feature == "pp_M",]$D,2)` , *p*
\< .05, it led to similar AUC than the standardized perimeter (*D* =
`r round(Results_ROCcomp[Results_ROCcomp$Feature == "Perimeter_zs",]$D,2)`,
*p* \< .05). This might be explained by the relatively lower sensitivity
`r All_ROC_round[All_ROC_round$Feature == "Polsby–Popper [score/9]",]$sensitivity`
(although relatively higher sensitivity
`r All_ROC_round[All_ROC_round$Feature == "Polsby–Popper [score/9]",]$specificity`)
of the Polsby-Popper test compared to the Perimeter
(`r All_ROC_round[All_ROC_round$Feature == "Perimeter [z-score]",]$sensitivity`)
and permuted validity tests
(`r All_ROC_round[All_ROC_round$Feature == "Permuted valid structure [boolean/9]",]$sensitivity`).
Additionally, the permuted validity led to significantly higher AUC than
the non permuted one (D =
`r round(Results_ROCcomp[Results_ROCcomp$Feature == "isValidStruct_M",]$D,2)`,
*p* \< .05).

Hence, the two best AUC performances are from the averaged permuted
validity score (AUC =
`r round(All_ROC[All_ROC$Feature == "isValid_perm_M",]$AUC,2)`; DP =
`r round(All_ROC[All_ROC$Feature == "isValid_perm_M",]$DP,2)`, cut-off =
`r round(All_ROC[All_ROC$Feature == "isValid_perm_M",]$threshold,2)`
corresponding to
`r round(All_ROC[All_ROC$Feature == "isValid_perm_M",]$threshold,2)*9`
valid of 9 forms) and the standardized perimeter between the repetitions
of each stimuli (AUC =
`r round(All_ROC[All_ROC$Feature == "Perimeter_zs",]$AUC,2)`; DP =
`r round(All_ROC[All_ROC$Feature == "Perimeter_zs",]$DP,2)`, cut-off =
`r round(All_ROC[All_ROC$Feature == "Perimeter_zs",]$threshold,2)`
z-scores). At the proposed cut-off, the permuted validity leads to
higher specificity than perimeter
(`r round(All_ROC[All_ROC$Feature == "isValid_perm_M",]$specificity,2)`
vs.
`r round(All_ROC[All_ROC$Feature == "Perimeter_zs",]$specificity,2)`)
indicating better rejection of controls (i.e. less false negatives), but
lower sensitivity
(`r round(All_ROC[All_ROC$Feature == "isValid_perm_M",]$sensitivity,2)`
vs.
`r round(All_ROC[All_ROC$Feature == "Perimeter_zs",]$sensitivity,2)`),
indicating poorer detection of SSS (i.e. more false negatives).

In sum, the permuted validity on the space-form level and the perimeter
of standardized coordinates on the stimulus level were the two equally
optimal tests to classify controls and SSS.

## Further analyses

We conducted further analyses by subsampling the dataset source and
questionnaire percentiles. The aim of these analyses was to test the
stability of the classifications from different methods across
experiments (i.e. datasets) and sample characteristics (i.e.
questionnaire scores). We also used a correlational approach to test for
correlations between features and questionnaires.

We sub-sampled the dataset source for the five most optimal tests (i.e.
higher AUC) and recalculated AUC, sensitivity and sensibility. We found
similar values for the different datasets, see @fig-myplotapx02.

Including only the data by Professor Ward, we subsampled the
participants depending on the perecntiles of the questionnaire scores
(hence only the data from Professor Ward), taking the 10-0% highest
questionnaire scores against the 10 % lowest (. (i.e. 90-100 %). Here
also most of the values remain relatively stable independently from the
questionnaire distribution of the samples, see @fig-myplotapx03. Another
important point addressed in the literature about consistency test is
circularity. The circularity is given in that if consistency tests are
used to classify SSS and controls, then those groups will by definition
differ in consistency [@root2025]. Hence, we made a correlation matrix
with the questionnaire scores and all the tests, see Appendix. The two
highest correlations are indeed between the questionnaire score and
perimeter (r = .58) and permuted validity (r = .50), see
@fig-myplotapx04.

Some of the participants might have been originally mis-attributed to
control or SSS or there might be consistent controls and inconsistent
synesthetes [see @root2025; @eagleman2007]. We found about \~6 % of the
full sample's synesthetes to be consistently classified as controls
(i.e. consistent false positive) by the five methods with the highest
AUC, see , see @fig-myplotapx07 and @fig-myplotapx08. \~3 % controls are
consistently classified as SSS (i.e. consistent false negative)
@fig-myplotapx09.

Finally, we also wanted to visualize the average space form of each
category at the population level to see if there were qualitative
differences between the forms. The figure in the appendix
(@fig-myplotapx06), shows more circular patterns for months and more
linear horizontal and vertical for numbers.

# Methods - future study {#sec-phase-ii-methods}

Additional data will be collected in the future using the same
consistency test, with the procedural exception that there will be four
repetitions per stimuli instead of three ( @sachdeva2024, Stage 1
in-principle acceptance). Hence for the stimulus levels tests we will
compare the area and perimeter of a quadrilateral of four coordinates
pairs instead of a triangle. Materials are more details on this study
are pre-registered on OSF, including ample size and recruitment method
[@sachdeva2024, Stage 1 in-principle acceptance] (see also
<https://osf.io/6wn4m>).

We will compute the same ROC analyses as in this pre-registration on all
the tests. Since we obtain different best AUC per tests depending on the
dataset (see @fig-myplotapx02) it is likely that the best test here will
not match with the best test in the to be collected dataset.

# Discussion - present study

Our investigation of four datasets found two main consistency tests
leading to the optimal classification of SSS from control: standardized
perimeter between stimulus repetitions and permuted topological validity
on the space-form level.

Further analyses show variabilities between the test's AUC exists
depending on the dataset source, which could explain why we could not
replicate @rothen2016's area but rather found the perimeter as a more
optimal test. The stability across varying percentiles of the
synesthesia questionnaire's score suggest the tests work as well for
more marked forms of SSS (note however that this analysis was limited
Professor Ward's data). The standardized perimeter also leads to
slightly higher correlation with the questionnaire scores (r = .57) than
permuted topological validity (r = .51), see @fig-myplotapx04.

In sum, while we could prove similar performances using a different
approach based on the space-form level (i.e. topological validity), we
found similar classification performances than using stimulus-repetition
level test such as perimeter.

## Limitations {#sec-limitations}

Although an optimal test to classify SSS might be particularly relevant
for experimental purposes, several important limitations need to be
considered. First, consistency tests contain only a limited set of
sequential stimuli (i.e. months, weeks and the first ten natural
numbers) that could potentially elicit synaesthetic experiences. Other
ordinal categories such as temperature, clock time, musical keys might
be more relevant for some individuals with SSS. For numbers
specifically, better consistency tests might be obtained using a larger
set size (i.e. including decades and hundreds), as descriptively
interesting form changes occur at different decimals in base-10 number
systems [@galton1880].

Second, the use of diagnostic cutoffs assumes categorical distinctions
between groups, but SSS may exist on a continuum and scores might be
more suitable [@price2013]. This limitation is related to the
circularity mentioned previously: diagnostic cutoffs as calculated here
depend on how SSS and controls are classified in the first place
[@simner2012]. Measures of other characteristics of synaesthesia such as
automaticity or visual Gabor detection might be necessary to establish
external validity [@ward2018]. Indeed, determining the prevalence of SSS
in the general population requires definitional choices that might be
based on more conservative or lenient criteria [@jonas2014; @brang2010;
@sagiv2006; @ward2018]. These difficulties are reflected in the span of
current prevalence estimates of SSS: 4.4 % [@brang2013], 8.1 %
[@ward2018] and 14 % [@seron1992].

Third, as shown in @fig-myplotapx02, we find that the optimal criteria
can vary across datasets. In other word, we obtain different AUC
depending on the datasets. These differences might be explained by
different sampling methods biases, for example [@vanpetersen2020]
recruited SSS from over one hundred candidates to maximise participant
reporting synesthetic experiences.

<!--#  Another possible explanation may be methodological and related to the option to skip responses in the task, leaving empty cases for some participants. Indeed, having fewer responses (i.e., fewer coordinates) increases the likelihood of producing a valid spatial form also in controls. Conversely, perimeter and area are differentially affected by missing responses. It is also possible that SSS are not actually inconsistent [@root2025; @eagleman2007] or that non self-reported SSS are consistent [@itoh2024]. Another possibility might have to do with the original classification of SSS. Indeed 55 % (127 of 231) of the controls from the data provided by Pr. Ward, report at least a space-form for Number Day or Month, see @fig-myplot1. Therefore some controls might also give consistent responses. Sequence-space associations could also be learned or memorized, leading to high consistency in non-synaesthetes as found for colour-grapheme [@ovallefresa2019].  -->

## Topological validity

Surprisingly, permuted topological validity of the spatial-forms led to
similar results than the perimeter test. This novel result might be
informative about how SSS map ordinal stimuli in space. Indeed, it seems
the patterns of spatial-forms follow topological rules analogous to
geographical and geometry-based space structures. Analogies between maps
and neuroscience have a long history [i.e. retinotopy, sonotopy or
somatotopy @eagleman2009]. Therefore spatial-forms might originate from
the mapping between ordinal or sequential stimuli on idiosyncratic
visuo-spatial abilities following topological rules. One of the
advantages of using topological validity is that it also classifies
responses with the same coordinates as invalid and hence inconsistent,
while using the area and perimeter metrics they would qualify as highly
consistent responses.

<!--#  Interestingly, ordinality is a very important semantic property of numbers, one of the categories used for SSS [@lyons2013]. Numbers are acquired sequentially (i.e. 1 →  2 →  3, ect. [@gelman1978] which would fit developmental accounts of the acquisition of SSS [@price2013].  -->

## Intermediary Conclusion {#sec-conclusions}

The present study compared traditional stimulus-level consistency tests
with novel form-level geometric tests for detecting SSS. We found
optimal classification performance from permuted topological validity
and standardized perimeter between repetitions. Despite not yielding to
better classification than stimulus level methods, this new method is
theoretically meaningful since genuine synesthetes should produce more
compact and topologically valid structures compared to controls.
Similarly, the perimeter test captures the stability of the overall
spatial configuration across repetitions, which should remain consistent
for true SSS individuals regardless of presentation order.

In the future study, we will attempt at validating these criteria on a
yet-to be acquired dataset.

# References

::: {#refs}
:::

{{< pagebreak >}}

### Appendix

The additional analyses in this appendix attempt to address several
concerns. Regarding the distribution's of the scores from the different
methods, i.e. whether it is bimodal, see @apx-featDensity. Then to
visualize the stability of the ROC analyses across the datasets, i.e.
whether AUC, DP, sensitivity and specificity vary across the four
different datasets, see @apx-SM2byds.

To address circularity concerns that we might find methods that best
classifies synesthetes and controls only from the sed on self-reported
groups, we attempted two additional approaches: first, testing
sub-samples based on percentiles of the questionnaire score sample
distribution (see @apx-byquestperc). The rationale is that a good
consistency test should have similar discriminative performance for
strong SSS and weaker SSS. To do this, we re-sampled the SSS and
controls based on their questionnaire distribution by percentiles (i.e.,
comparing 10% highest questionnaire scores vs. 10% lowest, then 20% vs.
20%, etc.). In addition, we correlated the questionnaire scores with the
results from different features extracted from the consistency test, see
@apx-correlation-with-self-report.

We also wanted to visually characterize whether certain patterns occur
more frequently for some categories than others (i.e. circular pattern
for months and linear for numbers). @apx-average-forms shows the average
forms for each categories. Finally, we wanted to visualize how each
participant is classified by each test. From this we found a group of
Synaesthetic who consistently are classified as controls by different
methods as well as a group of control who are consistently classified as
synaesthetes, see @apx-alluvial.

The following R packages were used for the data visualization in the
appendix:
`r pakret::pkrt_list("ggridges","cowplot","ggalluvial","corrplot")`

# Test's distributions {#apx-featDensity}

Regarding the distributions of each tests across the SSS vs. controls,
@fig-apx01 we compared the density plots of each standardized criteria
(in order to make them comparable) which visually suggests bimodal
distribution.

```{r}
#| label: fig-apx01
#| apa-note: 
#|    - Red = Classified as Synesthetes, Blue =  as controls. 
#|    - All method's score have been standardized (*z-score*). x axis treamed between -3 and 3 z-scores. Vertical line represent respectively 25 and 75 percentiles of each distributions.
#| fig-cap: Density plots of all the features comparing SSS and controls


ds_Q_zs <- ds_Q %>%
  select(feature_list, group,ID) %>%
  mutate_at(feature_list,scale) 

ds_Q_zs <- ds_Q_zs %>% 
  pivot_longer(cols = !c(group,ID),
               names_to = "Feature",
               values_to = "zs")


ds_Q_zs <- ds_Q_zs %>% 
  mutate(Feature = recode(Feature,
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" = "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "perm_zs" = "Permuted Chance level [z-score]",
                          "Self_Inter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" =  "Permuted valid structure [boolean/9]",
                          "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]"
                          ))

ds_Q_zs <- ds_Q_zs %>% mutate(group = recode(group,
                                            "Syn" = "Sequence-Space Synaesthetes",
                                             "Ctl" = "Controls"))

# So the features are sorted by AUC ;)
ds_Q_zs$Feature <- factor(ds_Q_zs$Feature, levels = rev(All_ROC_round$Feature))

ggplot(ds_Q_zs, aes(x = zs, y = Feature, fill = group, color  = group, point_color = group)) + 
  geom_density_ridges(alpha = 0.5, 
                      jittered_points = TRUE, 
                      position = position_points_jitter(width = 0.05, height = 0), 
                      point_shape = '|',
                      quantile_lines = TRUE, quantiles = c(0.25,0.75)) +
  xlim(-3, 3) +
  papaja::theme_apa() +
  scale_fill_manual(values = c("Sequence-Space Synaesthetes" = "steelblue",
                               "Controls" = "firebrick")) +
  scale_colour_manual(values = c("Sequence-Space Synaesthetes" = "steelblue",
                               "Controls" = "firebrick")) +
  labs(y = "Method", x = "z-score") +
    theme(legend.position = "none")
```

```{r}
#| label: tbl-mytbapx01
#| eval: false
#| ft-align: left
#| include: false
#| tbl-cap: Median and SD of each standardized features

feat_cols <- colnames(ds_Q[,42:length(colnames(ds_Q))])

fmt_ms <- function(x, digits = 2) {
  m    <- mean(x, na.rm = TRUE)
  Med  <- median(x, na.rm = TRUE)
  sd   <- stats::sd(x, na.rm = TRUE)
  paste0(# "M = ",
  #        formatC(m,  format = "f", digits = digits),
         " Med = ",
         formatC(Med,  format = "f", digits = digits),
         " (",
         formatC(sd, format = "f", digits = digits),
         ")"
  )
}

tmp <- ds_Q_zs %>%
    group_by(group,  Feature) %>%
    summarise(`M (SD)` = fmt_ms(zs), .groups = "drop") %>%
    pivot_wider(names_from = c(group),
                values_from = c(`M (SD)`)) 

tmp <- tmp[match(All_ROC_round$Feature, tmp$Feature), ]
# tmp <- tmp %>% rename("Control" = "Ctl", "Sequence-Space Synaesthetes" = "Syn")

apa_table(tmp)
rm(tmp)
```

# ROC analysis by dataset {#apx-SM2byds}

Here we compare the ROC analyses results for each data sample of the
three methods with the best AUC, see @fig-apx08. Descriptively, we also
looked a the distributions of each methods across SSS and controls
@fig-myplotapx02. Note that the dataset labelled as Ward 2 has more
synaesthetes than controls (5:1 ratio). Note that some of the tests led
to 100 % Specificity in @vanpetersen2020 dataset.

```{r}
#| label: tbl-mytbapx02
#| eval: false
#| include: false
#| tbl-cap: Average feature for Synesthetes by dataset

feat_cols <- All_ROC$Feature[order(All_ROC$AUC, decreasing = TRUE)]

tbl_out <- ds_Q %>%
    # filter(dataSource != "Ward2") %>%
    select(c(group, dataSource, feat_cols)) %>%
    pivot_longer(cols = all_of(feat_cols),
                 names_to = "Feature",
                 values_to = "Value") %>%
    group_by(group, dataSource, Feature) %>%
    # summarise(`M (SD)` = fmt_ms(Value), .groups = "drop") %>%
    summarise(Median = median(Value), .groups = "drop") %>%
    pivot_wider(names_from = c(group, dataSource),
                values_from = Median) %>%
    arrange(match(Feature, feat_cols))

tbl_out <- tbl_out %>% rename("vanPeters " = "Ctl_PeterCor", 
                   "Rothen " = "Ctl_Rothen", 
                   "Ward " = "Ctl_Ward", 
                   "Ward2 " = "Ctl_Ward2",
                   "vanPeters" = "Syn_PeterCor", 
                   "Rothen" = "Syn_Rothen", 
                   "Ward" = "Syn_Ward", 
                   "Ward2" = "Syn_Ward2")

tbl_out <- tbl_out %>% mutate(Feature = recode(Feature,
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" = "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "perm_zs" = "Permuted Chance level [z-score]",
                          "Self_Inter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" =  "Permuted valid structure [boolean/9]",
                           "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]"))

# I struggled so much with those tables, that I decided to have a density plot. 
# Looks nicer and more informative. 

# knitr::kable(tbl_out[,c(1,6:9)],format = "latex", booktabs = TRUE, digits = 3)
# knitr::kable(tbl_out[,c(1,6:9)],format = "latex", booktabs = TRUE, digits = 3)
# tbl_out
# papaja::apa_table(tbl_out,
#       col_spanners = list("Controls" = c(2,5), "Sequence-Space Synaesthetes" = c(6,9))) # "Currently ignored in Word documents."
# knitr::kable(tbl_out[,1:5],digits = 3)
# tbl_out[,1:5] %>%
#   flextable::flextable() %>% 
#   flextable::theme_apa()
```

```{r AUCbyDataSource}

dataSources <- unique(ds$dataSource)
AUC_ds <- vector("list", length(dataSources))


for(i in 1:length(dataSources)){
  
  ds_here <- ds_Q %>% filter(dataSource %in% dataSources[i])
  
  # INITIALIZE ROC_curves
  ROC_curves <- list()
  ROC_contingency <- list()
  
  # Initialize dataframe to collect relevant ROC values:
  Per_ds_ROC <- data.frame(Feature=character(),
                        AUC = character(),
                        threshold=integer(),
                        sensitivity=integer(),
                        specificity=integer(),
                        ppv = integer(),
                        npv = integer(),
                        high_ci = integer(),
                        low_ci = integer(),
                        stringsAsFactors=FALSE)
  
  # Loop by features:
  for(fit_n in 1:length(Feats[,1])){
    ROC_here <- Comp_ROC(ds_here, "group", Feats[[fit_n,1]],"ID", Feats[[fit_n,2]])
    
    ROC_curves      <- c(ROC_curves,list(ROC_here$ROC_curve,ROC_here$ROC_properties$Feature))
    Per_ds_ROC      <- rbind(Per_ds_ROC, ROC_here$ROC_properties)
    ROC_contingency <- c(ROC_contingency,list(ROC_here$Coningency_table,
                                              ROC_here$ROC_properties$Feature))
  }
  
  Per_ds_round <- Per_ds_ROC %>% 
    mutate_if(is.numeric, round,2)
  
  Per_ds_round <- Per_ds_round[order(Per_ds_round$AUC, decreasing = TRUE), ]
  
  
   AUC_ds[[i]] <- data.frame(
    Feature = Per_ds_round$Feature,
    AUC = Per_ds_round$AUC,
    DP =  Per_ds_round$DP,
    Sensitivity = Per_ds_round$sensitivity,
    Specificity = Per_ds_round$specificity,
    dataSource = dataSources[i],
    N = length(ds_here$ID)
  )
}

AUC_ds <- do.call(rbind, AUC_ds)

# Comment the following if you want to see all features. I thought the figure were a bit overwhelming for my tiny cognitive capacities
AUC_ds <- AUC_ds %>% 
  #filter(dataSource != "Ward2") %>%
  mutate(Feature = recode(Feature,
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" = "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "perm_zs" = "Permuted Chance level [z-score]",
                          "Self_Inter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" =  "Permuted valid structure [boolean/9]",
                           "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]")) %>%
  
  filter(Feature %in%  All_ROC_round$Feature[1:5]) %>%
  mutate(dataSource = recode(dataSource,"PeterCor" = "vanPetersen")) 

```

```{r}
#| label: fig-myplotapx02
#| apa-note: "AUC = Area Under the Curve, DP = Discrimination Power"
#| fig-cap: "Lineplots of AUC and DP by data source"

AUC_ds$dataSource <- as.factor(AUC_ds$dataSource)
AUC_ds <- AUC_ds %>% 
  filter(dataSource != "Ward2") %>%
   mutate(dataSource = factor(dataSource, levels=c("Rothen", "Ward", "vanPetersen")))

gp_AUC <- ggplot(AUC_ds, aes(x = dataSource, y = AUC, group = Feature, colour = Feature)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values =pals::cols25(nFeat)) + 
  papaja::theme_apa() +
   xlab('') +
  ylim(50,100) +
  theme(legend.position="none",axis.text.x = element_blank())

gp_DP <- ggplot(AUC_ds, aes(x = dataSource, y = DP, group = Feature, colour = Feature)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values =pals::cols25(nFeat)) +
  papaja::theme_apa() +
  xlab(' ') +
  ylim(0,6) +
  theme(legend.position="none",axis.text.x = element_blank())

gp_Sensitivity <- ggplot(AUC_ds, aes(x = dataSource, y = Sensitivity, group = Feature, colour = Feature)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values =pals::cols25(nFeat)) + 
  papaja::theme_apa() +
  xlab(' ') +
  ylim(50,100) +
  theme(legend.position="none",axis.text.x = element_text(angle = 90))

gp_Specificity <- ggplot(AUC_ds, aes(x = dataSource, y = Specificity, group = Feature, colour = Feature)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values =pals::cols25(nFeat)) + 
  papaja::theme_apa() +
  xlab(' ') +
  ylim(50,100) +
  theme(legend.position="none",axis.text.x = element_text(angle = 90))


legend <- get_legend(
  ggplot(AUC_ds, aes(x = dataSource, y = Specificity, group = Feature, colour = Feature)) +
    geom_point() +
    papaja::theme_apa() +
    scale_color_manual(values =pals::cols25(nFeat)) +
    guides(color = guide_legend(nrow = 2)) +
    theme( legend.justification = c(0.2,1), legend.position = "bottom",
           legend.text=ggplot2::element_text(size=10))
)


plot_grid(gp_AUC, gp_DP,gp_Sensitivity,gp_Specificity, 
          labels = c("A","B","D","E"),legend = legend, ncol=2,
          rel_heights = c(1, 1.3, 0.4)) # height of each row. 3rd row is legend, 10% of the others.
```

```{r}
#| label: fig-apx08
#| apa-note: 
#|    - Red = Classified as Synesthetes, Blue =  as controls. 
#|    - All method's score have been standardized (*z-score*). x axis treamed between -3 and 3 z-scores. Vertical line indicate respectively 25 and 75 percentiles of each distributions.
#| fig-cap: Density plots of all the tests comparing SSS and controls by dataset


ds_Q_zs <- ds_Q %>%
  select(feature_list, group,ID, dataSource) %>%
  mutate_at(feature_list,scale) 

ds_Q_zs <- ds_Q_zs %>% 
  pivot_longer(cols = !c(group,ID, dataSource),
               names_to = "Feature",
               values_to = "zs")


ds_Q_zs <- ds_Q_zs %>% 
  mutate(Feature = recode(Feature,
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" = "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "perm_zs" = "Permuted Chance level [z-score]",
                          "Self_Inter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" =  "Permuted valid structure [boolean/9]",
                          "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]"
                          ))

ds_Q_zs <- ds_Q_zs %>% mutate(group = recode(group,
                                            "Syn" = "Sequence-Space Synaesthetes",
                                             "Ctl" = "Controls"))

# So the features are sorted by AUC ;)
ds_Q_zs$Feature <- factor(ds_Q_zs$Feature, levels = rev(All_ROC_round$Feature))

ggplot(ds_Q_zs, aes(x = zs, y = Feature, fill = group, color  = group, point_color = group)) + 
  geom_density_ridges(alpha = 0.5, 
                      jittered_points = TRUE, 
                      position = position_points_jitter(width = 0.05, height = 0), 
                      point_shape = '|',
                      quantile_lines = TRUE, quantiles = c(0.25,0.75)) +
  xlim(-3, 3) +
  papaja::theme_apa() +
  scale_fill_manual(values = c("Sequence-Space Synaesthetes" = "steelblue",
                               "Controls" = "firebrick")) +
  scale_colour_manual(values = c("Sequence-Space Synaesthetes" = "steelblue",
                               "Controls" = "firebrick")) +
  labs(y = "Method", x = "z-score") +
  facet_grid(~ dataSource) +
    theme(legend.position = "none",axis.text=element_text(size=9))
```

# ROC Analysis Sub-sampled data by questionnaire quantiles (20% steps) {#apx-byquestperc}

We compared the data sampled by the questionnaire score. Based on the
distribution of the questionnaire score, we sampled the 10 % with the
lowest and 10 % with the highest scores. Those are then compared with
the 20 and 20 % and so on until 40 and 40 %. The rationale of this
procedure is that AUC, sensitivity and specificity should remain stable
across percentiles for a feature to be valid, see @fig-myplotapx03. In
other words the ROC should remain unchanged if we take extreme groups
compared to less extreme ones.

```{r}
ds_Q2 <- ds_Q %>% filter(dataSource %in% c("Ward","Ward2")) %>% filter(!is.na(`questionnaire score`))

p <- seq(0,1,0.10)
quant <- quantile(ds_Q2$`questionnaire score`, probs = p)
quant <- as.numeric(quant)

AUC_perctils <- vector("list", length(quant)/2)


for(quant_n in 1:(length(quant)/2)){
  
  All_ROC <- data.frame(Feature=character(),
                        AUC = character(),
                        threshold=integer(),
                        sensitivity=integer(),
                        specificity=integer(),
                        ppv = integer(),
                        npv = integer(),
                        high_ci = integer(),
                        low_ci = integer(),
                        stringsAsFactors=FALSE)
  
  ds_Q2_quant_n <- ds_Q2 %>% 
  filter(`questionnaire score` <= quant[quant_n] | `questionnaire score` >= quant[length(quant)-quant_n])
  
  # hist(ds_Q2_quant_n$`questionnaire score`)
  
  # Loop by features:
  for(fit_n in 1:length(Feats[,1])){
    ROC_here <- Comp_ROC(ds_Q2_quant_n, "group", Feats[[fit_n,1]],"ID",Feats[[fit_n,2]])
    
    ROC_curves      <- c(ROC_curves,list(ROC_here$ROC_curve,ROC_here$ROC_properties$Feature))
    All_ROC         <- rbind(All_ROC, ROC_here$ROC_properties)
    ROC_contingency <- c(ROC_contingency,list(ROC_here$Coningency_table,ROC_here$ROC_properties$Feature))
  }
  
  All_ROC_round <- All_ROC %>% 
    mutate_if(is.numeric, round,2)
  All_ROC_round <- All_ROC_round[order(All_ROC_round$AUC, decreasing = TRUE), ]
  
  AUC_perctils[[quant_n]] <- data.frame(
    Feature = All_ROC_round$Feature,
    AUC = All_ROC_round$AUC,
    DP = All_ROC_round$DP,
    Sensitivity = All_ROC_round$sensitivity,
    Specificity = All_ROC_round$specificity,
    Quant_min = quant[quant_n],
    Quant_max = quant[length(quant)-quant_n],
    quant_n = quant_n
  )

    # print(knitr::kable(All_ROC_round))
}

AUC_perctils <- do.call(rbind, AUC_perctils)

AUC_perctils$quant_n <- as.factor(AUC_perctils$quant_n)
levels(AUC_perctils$quant_n) <- c("1" = "10 & 90%",
                          "2" = "20 & 80%",
                          "3" = "30 & 70%",
                          "4" = "40 & 60%",
                          "5" = "100%")
AUC_perctils <- AUC_perctils %>%
  filter(Feature %in%  All_ROC$Feature[order(All_ROC$AUC, decreasing = TRUE)][1:5]) %>%
  mutate(Feature = recode(Feature,
                          "Area_perc" = "Area [screen size %]",
                          "Area_zs" = "Area [z-score]", 
                          "Perimeter_perc" = "Perimeter [screen size %]",
                          "Perimeter_zs" = "Perimeter [z-score]",
                          "perm_zs" = "Permuted Chance level [z-score]",
                          "Self_Inter" = "Segment self-intersections [n/9]",
                          "areaPoly_GA" = "Area of the polygon [z-score]",
                          "perimPoly" = "Perimeter of the polygon [z-score]" ,
                          "isSimple_GA" = "Simple topology [boolean/9]",
                          "SelfInter" = "Segment self-intersections [n/9]",
                          "isValidStruct_M" = "Valid structure [boolean/9]",
                          "isValid_perm_M" =  "Permuted valid structure [boolean/9]",
                           "pp_M" = "Polsby–Popper [score/9]",
                          "pp_perm_M" = "Permuted P-P [score/9]"))
  
```

```{r}
#| label: fig-myplotapx03
#| fig-cap: "Lineplots of AUC and DP by questionnaire percentile"
#| apa-note: "Each point is an increasing percentiles. AUC = Area Under the Curve. DP = Discrimination Power. Specitficites of some tests with van Petersen data are 100 %, so the DP are nor computable nor interpretable"

gp_AUC <- ggplot(AUC_perctils, aes(x = quant_n, y = AUC, group = Feature, colour = Feature)) +
  geom_path() +
  geom_point() +
  scale_color_manual(values =pals::cols25(nFeat)) + #
  papaja::theme_apa() +
  xlab(' ') +
  ylim(50,100) +
  theme(legend.position="none",axis.text.x = element_blank())

gp_DP <- ggplot(AUC_perctils, aes(x = quant_n, y = DP, group = Feature, colour = Feature)) +
  geom_path() +
  geom_point() +
  scale_color_manual(values =pals::cols25(nFeat)) + #
  papaja::theme_apa() +
  xlab(' ') +
  ylim(0,6) +
  theme(legend.position="none",axis.text.x = element_blank())

gp_Sensitivity <- ggplot(AUC_perctils, aes(x = quant_n, y = Sensitivity, group = Feature, colour = Feature)) +
  geom_path() +
  geom_point() +
  scale_color_manual(values =pals::cols25(nFeat)) + #
  papaja::theme_apa() +
  xlab(' ') +
  ylim(50,100) +
  theme(legend.position="none",axis.text.x = element_text(angle = 90))

gp_Specificity <- ggplot(AUC_perctils, aes(x = quant_n, y = Specificity, group = Feature, colour = Feature)) +
  geom_path() +
  geom_point() +
  scale_color_manual(values =pals::cols25(nFeat)) + #
  papaja::theme_apa() +
  xlab(' ') +
  ylim(50,100) +
  theme(legend.position="none",axis.text.x = element_text(angle = 90))


legend <- get_legend(
  ggplot(AUC_perctils, aes(x = quant_n, y = Specificity, group = Feature, colour = Feature)) +
    geom_point() +
    papaja::theme_apa() +
    scale_color_manual(values =pals::cols25(nFeat)) +
     guides(color = guide_legend(nrow = 2)) +
    theme( legend.justification = c(0.2,1), legend.position = "bottom",
           legend.text=ggplot2::element_text(size=10))
)


plot_grid(gp_AUC, gp_DP,gp_Sensitivity,gp_Specificity, 
          labels = c("A","B","D","E"),legend = legend, ncol=2,
          rel_heights = c(1, 1.3, 0.4)) # height of each row. 3rd row is legend, 10% of the others.
```

# Correlation with self-report {#apx-correlation-with-self-report}

The best criterion should also best correlate with SSS self-reported
questionnaire score. Works only with Ward's aggregated data, see
@fig-myplotapx04.

```{r}
#| label: fig-myplotapx04
#| message: false
#| warning: false
#| apa-note: Only data from Ward is included here
#| fig-cap: Correlation with self-reported questionnaire
#| fig-height: 9
#| fig-width: 9

# Or shorter version (no details on questionaire):
sel_dsQ <- ds_Q %>%
  select(c(`questionnaire score`,feature_list)) %>%
  filter(!is.na(`questionnaire score`))  %>%
  select_if(~ !any(is.na(.)))

cols <- colnames(sel_dsQ)
cols[cols %in% SortFeat] <- SortFeat

sel_dsQ <- sel_dsQ[, cols]

sel_dsQ <- sel_dsQ %>% 
  select(c("questionnaire score",All_ROC_round$Feature[1:5])) %>%
  rename("Perimeter [z-score]" = "Perimeter_zs",
         "Valid structure [mean binary/9]" = "isValidStruct_M",
         "Permuted valid structure [mean binary/9]" = "isValid_perm_M",
         "Polsby–Popper [mean(score)]" = "pp_M",
          "Permuted P−P [score/9]" = "pp_perm_M")

sel_CorMat <- cor(sel_dsQ)
sel_CorMat_test <- cor.mtest(sel_dsQ)
  

corrplot(abs(sel_CorMat), method = 'number', diag=FALSE, type = "lower",tl.col = 'black', cl.pos = 'n',col = c('black', 'black'))

```

# Centred average forms {#apx-average-forms}

In this additional analyses we attempted at visualizing the average for
each categories to see whether a pattern could be discerned on a
qualitative level. To avoid negative coordinates, we shifted all the
z-scores from the whole dataset minimum for the x and y coordinates.
Then we anchored to the coordinate 0,0 the stimulus "1" for numbers,
"Monday" for weekdays and "January" for months. We then plotted all the
space forms with transparency, see @fig-myplotapx06.

Descriptively it is possible to see more circular pattern for months and
horizontal and vertical patterns for numbers. Also the averaged forms
appear to be more clearly defined in the synaesthetes than control.

```{r}

# 1. I need to shift all the data to avoid negative numbers 
ds$x_zs_shift <- ds$x_zs -min(ds$x_zs) # Can also run with x and y.
ds$y_zs_shift <- ds$y_zs -min(ds$y_zs)

# 2. Generate the vector with the to be substracted values
target_stimuli <- c("1", "Monday", "January")

# Extract the target stimuli values
target_values <- ds %>%
  filter(stimulus %in% target_stimuli) %>%
  select(ID, Cond, repetition, stimulus, x_zs_shift, y_zs_shift)

# anchor each figure
ds_adj <- ds %>%
  left_join(
    target_values %>%
      pivot_wider(
        names_from = stimulus,
        values_from = c(x_zs_shift, y_zs_shift),
        names_sep = "_stim_"
      ),
    by = c("ID", "Cond", "repetition")
  )

# Adjust the coordinates based on the anchor by categories:
# x coordinate
## Numbers
ds_adj$x_adj[!is.na(ds_adj$x_zs_shift_stim_1)] <- 
  ds$x_zs_shift[!is.na(ds_adj$x_zs_shift_stim_1)] - 
  ds_adj$x_zs_shift_stim_1[!is.na(ds_adj$x_zs_shift_stim_1)]
## Weekdays
ds_adj$x_adj[!is.na(ds_adj$x_zs_shift_stim_Monday)] <- 
  ds$x_zs_shift[!is.na(ds_adj$x_zs_shift_stim_Monday)]  - 
  ds_adj$x_zs_shift_stim_Monday[!is.na(ds_adj$x_zs_shift_stim_Monday)]
## Month
ds_adj$x_adj[!is.na(ds_adj$x_zs_shift_stim_January)] <- 
  ds$x_zs_shift[!is.na(ds_adj$x_zs_shift_stim_January)] - 
  ds_adj$x_zs_shift_stim_January[!is.na(ds_adj$x_zs_shift_stim_January)] 

# same for y coordinate
ds_adj$y_adj[!is.na(ds_adj$y_zs_shift_stim_1)] <- 
  ds$y_zs_shift[!is.na(ds_adj$y_zs_shift_stim_1)]  - 
  ds_adj$y_zs_shift_stim_1[!is.na(ds_adj$y_zs_shift_stim_1)] 

ds_adj$y_adj[!is.na(ds_adj$y_zs_shift_stim_Monday)]  <- 
  ds$y_zs_shift[!is.na(ds_adj$y_zs_shift_stim_Monday)]  - 
  ds_adj$y_zs_shift_stim_Monday[!is.na(ds_adj$y_zs_shift_stim_Monday)] 

ds_adj$y_adj[!is.na(ds_adj$y_zs_shift_stim_January)]  <- 
  ds$y_zs_shift[!is.na(ds_adj$y_zs_shift_stim_January)]  - 
  ds_adj$y_zs_shift_stim_January[!is.na(ds_adj$y_zs_shift_stim_January)] 

```

```{r}
#| label: fig-myplotapx06
#| fig-cap: "Average Centered space forms visualized"
#| fig-format: png
#| dpi: 300
#| apa-notes: "Average of all the space-forms with centers 0,0 for January, 1 and Monday respectively"

ds_adj$group <- recode(ds_adj$group, "Ctl" = "Control", "Syn" = "SSS")

# AVERAGE FIGURE:
ds_adj %>% 
  # filter(ID %in% "84735") %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ggplot(aes(x = x_adj, y = y_adj, group = as.factor(repetition))) +
  geom_path(aes(x = x_adj, y = y_adj), alpha = 0.02, colour = "black") +
  facet_wrap(group ~ Cond) +
  papaja::theme_apa() +
  labs(x = "x [ centered z-score]",
       y = "y [ centered z-score]")

```

# Alluvial plot {#apx-alluvial}

Are false positives and negatives share some characteristics? To answer
this question qualitatively we identified two group of self-reported SSS
participants who care consistently classified as controls by the five
methods with the highest AUC (i.e. consistent false negatives), see dark
blue lines in @fig-myplotapx07. On the other side we identified a group
of controls who are consistently classified as SSS by the five methods
with the highest AUC, see dark red lines in @fig-myplotapx07 .

```{r}
#| label: fig-myplotapx07
#| fig-cap: "Alluvial plot of test passing"
#| apa-note: "Red = Controls, Blue = Synaesthetes. Highlighted in blue are Synesthetes who are classified as control by all 5 tests. Highlighted in red are controls who are classified as synesthetes by all 5 tests."



AllList       <- list(Group = ds_Q$ID[ds_Q$group == "Syn"])
ds_Q$pass_SynGroup <- ds_Q$group == "Syn"

All_ROC_round <- All_ROC %>% 
  mutate_if(is.numeric, round,2)
All_ROC_round <- All_ROC_round[order(All_ROC_round$AUC, decreasing = TRUE), ]

for(feat_i in 1:length(All_ROC_round$Feature) ){
  feat_here <- All_ROC_round$Feature[feat_i] 
  
  if(feat_i %in% c(2,3,4,5)){ # Sorry for the lazy programming here 
    
    ds_Q[[paste0("pass_",feat_here)]]  <- ds_Q[[feat_here]] >= All_ROC_round$threshold[feat_i]
    AllList[[feat_here]] <- (ds_Q$ID[ds_Q[[feat_here]] >= All_ROC_round$threshold[feat_i]])
    
  } else {
    
    ds_Q[[paste0("pass_",feat_here)]]  <- ds_Q[[feat_here]] <= All_ROC_round$threshold[feat_i]
    AllList[[feat_here]] <- (ds_Q$ID[ds_Q[[feat_here]] <= All_ROC_round$threshold[feat_i]])
  }
}

df_long <- ds_Q %>%
  filter(!ID %in%  "69253") %>%
  select("ID", starts_with("pass")) %>%
  mutate( ID = factor(ID),
         Syngroup = factor(pass_SynGroup)) %>%
  pivot_longer(
    cols = starts_with("pass"),
    names_to = "test",
    values_to = "result"
  )


test_order <- c("pass_SynGroup", paste0("pass_",All_ROC_round$Feature))
df_long$test <- factor(df_long$test, levels = test_order)

df_long <- df_long %>% 
  filter(test %in% levels(df_long$test)[1:6]) %>%
  mutate(test = recode(test, "pass_SynGroup" = "Group",
                       "pass_Perimeter_zs" = "Perimeter [z-score]",
                       "pass_isValidStruct_M" = "Valid structure [mean binary/9]",
                       "pass_Area_zs" = "Area [z-score]", 
                       "pass_isValid_perm_M" =  "Permuted valid structure [mean binary/9]",
                       "pass_SelfInter" = "Segment self-intersections [n/9]",
                       "pass_perm_zs" = "Permuted Chance level [z-score]",
                       "pass_pp_M" = "Polsby–Popper [score/9]"))


## ID that fail all 5 tests (except the group one) among the controls
df_long$test <- as.factor(df_long$test)

ID_SynFail5Test <- df_long %>% 
  filter(Syngroup %in% TRUE) %>% # Only synesthetes (TRUE = Syn)
  filter(test %in% levels(df_long$test)[2:6]) %>%  # select first 5 tests
  group_by(ID) %>% 
  mutate(totRes = sum(result)) %>%
  filter(totRes == 0) # 1 because we also have the Snygroup

ID_SynFail5Test <- unique(ID_SynFail5Test$ID)

## ID Ctl pass all tests (except the group one) 
ID_CtlPass5Test <- df_long %>% 
  filter(Syngroup %in% FALSE) %>% # Only controls 
  filter(test %in% levels(df_long$test)[2:6]) %>%  # select first 5 tests
  group_by(ID) %>% 
  mutate(totRes = sum(result)) %>%
  filter(totRes == 5) # 1 because we also have the Snygroup

ID_CtlPass5Test <- unique(ID_CtlPass5Test$ID)

df_long$Syngroup <- as.character(df_long$Syngroup)
df_long$Syngroup[df_long$ID %in% ID_SynFail5Test] <- "Highlight"
df_long$Syngroup[df_long$ID %in% ID_CtlPass5Test] <- "Highlight2"

df_long %>%
  filter(test %in% levels(df_long$test)[1:6]) %>%# Keep top10 for readability  
  # mutate(result = !result) %>% # Hence true = Syn
  ggplot(aes(x = test, stratum = factor(result),
             alluvium = ID, # Keep the same color for the same ID across tests
             fill = Syngroup, label = result)) +
  geom_flow(stat = "alluvium",alpha = 1) +
  geom_stratum(aes(fill = factor(result)), alpha = 0.6) +
  scale_fill_manual(values = c("TRUE" = "steelblue",
                               "FALSE" = "firebrick", 
                               "Highlight" ="blue",
                                "Highlight2" ="red")) +
  theme_minimal() +
  labs(x = "Test",
       y = "Number of IDs") +
  theme(axis.ticks.x=element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90),
        legend.position = "none")
```

On one side
`r round(length(unique(ID_SynFail5Test))/length(ds_Q$ID)*100,2)` % of
all participants, that is `r length(unique(ID_SynFail5Test))`
Synaesthetes where consistently classified as controls by all five
tests, see @fig-myplotapx08. On the other side
`r round(length(unique(ID_CtlPass5Test))/length(ds_Q$ID)*100,2)` % of
all participants, `r length(unique(ID_CtlPass5Test))` controls where
consistently classified as SSS by all five tests, see @fig-myplotapx09.
Given the different approaches of each methods, the identified
participants might be on one side non-genuine synesthetes and on the
other controls who have SSS but are either not aware or did not identify
it as such.

```{r}
#| label: fig-myplotapx08
#| fig-cap: Synaesthetes who consistently *fail* all five tests (blue in the alluvial plot)

ds %>% 
  filter(ID  %in% ID_SynFail5Test) %>%
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>% # Start ggplot
  ggplot(aes(x = x_zs, y = y_zs, group = stimulus, label = stimulus, fill = stimulus)) +
  # geom_text(aes(x = X_mean_zs+0.1, y = Y_mean_zs+0.1), colour = "black", size = 0.5) +
  # geom_path(aes(x = X_mean_zs, y = Y_mean_zs, group = 1)) +
  geom_path(aes(x = x_zs, y = y_zs, group = repetition), alpha = 0.2) +
  geom_text(aes(x = x_zs+0.1, y = y_zs+0.1), size = 0.5, alpha = 0.5) +
  facet_wrap(ID ~ Cond) +
  theme_void()  +
  theme(strip.text = element_text(size = 5))

```

```{r}
#| label: fig-myplotapx09
#| fig-cap: Controls who consistently *pass* all five tests (red in the alluvial plot)
ds %>% 
  filter(ID  %in% ID_CtlPass5Test) %>%
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>% # Start ggplot
  ggplot(aes(x = x_zs, y = y_zs, group = stimulus, label = stimulus, fill = stimulus)) +
  # geom_text(aes(x = X_mean_zs+0.1, y = Y_mean_zs+0.1), colour = "black", size = 0.5) +
  # geom_path(aes(x = X_mean_zs, y = Y_mean_zs, group = 1)) +
  geom_path(aes(x = x_zs, y = y_zs, group = repetition), alpha = 0.2) +
  geom_text(aes(x = x_zs+0.1, y = y_zs+0.1), size = 0.5, alpha = 0.5) +
  facet_wrap(ID ~ Cond) +
  theme_void() +
  theme(strip.text = element_text(size = 5))


# row1 <- ds_Q %>% 
#   filter(ID %in% ID_SynFail5Test) %>%
#   select(starts_with("Q")) %>%
#   summarise(across(everything(), mean, na.rm=TRUE))
# 
# row2 <- ds_Q %>% 
#     filter(group %in% "Syn") %>%
#     select(starts_with("Q")) %>%
#     summarise(across(everything(), mean, na.rm=TRUE))
# 
# bind_rows(row1,row2)
```

````{=html}
<!--# 

# Code to visualize all spatial-forms

This exports many pdf's. It plots each ID and condition z-score x and y
coordinates. Since each coordinate is repeated 3 times, these are
represented by triangles. The line paths connect average coordinates to
visualize forms (stimulus are ordered, i.e. 1 to 9, Monday to Sunday,
January to December). Finally in the top right corner, each dots
indicates if the ID would pass (green dot) / fails (red dot) depending
on the criteria.

```{r}
# Match ds and ds_Q criteria

ds <- left_join(ds, 
          ds_Q %>% select("ID",starts_with("pass")), 
          by = "ID")

# To add a form by the middle
ds <- ds %>% group_by(ID,stimulus) %>% mutate(X_mean_zs = mean(x_zs), Y_mean_zs = mean(y_zs))

ds$passGroup <- ds$group == "Syn"
```

```{r}
#| eval: false
#| include: false
# Multiple pages


IDlist <- unique(ds$ID)

# Set the number of rows & columns per pages
N_rows = 5
N_cols = 9

for(i in 1:(round(length(IDlist)/(N_rows*N_cols)*3)+1)){
  
 ds %>% 
    # filter(ID  %in% c(28779,29027,29043)) %>%
    group_by(stimulus) %>%
    arrange(stimulus) %>%
    arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>% # Start ggplot
    ggplot(aes(x = x_zs, y = y_zs, group = stimulus, label = stimulus, fill = stimulus)) +
    geom_text(aes(x = 4, y = 4.5, label = dataSource), size = 1) + 
    geom_text(aes(x = 3.7, y = 5.5, label = c("Criteria:  Ques   Perim  Area    Area_zs     validPerm")), size = 2) + 
    geom_point(aes(x = 5 -2, y = 5, color = passGroup), size = 0.5) + 
    geom_point(aes(x = 5 -1.5, y = 5, color =pass_Perimeter_zs), size = 0.5) + 
    geom_point(aes(x = 5 -1., y = 5, color =pass_Area), size = 0.5) +
    geom_point(aes(x = 5 -0.5, y = 5, color =pass_Area_zs), size = 0.5) +
    geom_point(aes(x = 5 + 0.2, y = 5, color =pass_isValid_perm_M), size = 0.5) + # SynLine is developed later in the code.
    geom_polygon(alpha = 0.4) +
    geom_text(aes(x = X_mean_zs+0.1, y = Y_mean_zs+0.1), colour = "black", size = 0.5) +
    geom_path(aes(x = X_mean_zs, y = Y_mean_zs, group = 1)) +
    geom_path(aes(x = x_zs, y = y_zs, group = repetition), alpha = 0.2) +
    geom_text(aes(x = x_zs+0.1, y = y_zs+0.1), size = 0.5, alpha = 0.5) +
    facet_wrap_paginate( ~ ID+ Cond, ncol = N_cols, nrow = N_rows, page = i)  +
    theme_minimal()
    ggsave(paste0("Figures_SM/Syn_Categories",i,".pdf"),width = N_cols*2, height = N_rows*2)
}
```

 -->
````
