---
title: "Pre-registered report: Space Sequence Synesthesia Diagnostic using form mapping"
output:
  html_document: default
date: today
editor_options:
  markdown:
    wrap: 72
number-sections: true
link-citations: true
bibliography: references.bib
---

```{r}
#| eval: false
#| include: false

format: pptx
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
```

```{r, setup, include=FALSE}
#   
knitr::opts_chunk$set(
  fig.width = 10, fig.height = 8, echo = F
)

# Ev. if have all the data pre-saved (shortens knitting, especially the permutation take long)
# load("DataSave3S_V5.RData")
```

```{r, libraries, message=FALSE, warning=FALSE}
library(readr)
library(readxl)

library(tidyr)
library(dplyr)

library(papaja)

library(ggplot2)
library(ggridges)
library(ggalluvial)

library(pROC) # See https://www.r-bloggers.com/2019/02/some-r-packages-for-roc-curves/
```

# Abstract:

Existent diagnostic tools for space sequence synethesia are based on
questionnaire and response consistency. Consistency is calculated as the
area between repetitions for the same inducer. In the first present
phase, available data from 467 participants is used to explore new
geometrical features to discriminate syntheses from controls.
Conceptually, our goal is to take advantage of the inducer's ordinality
that create synesthetic forms. For this aim, we harness a geography
package to extract geometrical features to use as a test for synethesia.
Reciever Operator Characteristics are used to select the features that
best diagnose synesthesia. In a second phase to come, we test the
predictive power of the new diagnostic features onto newly collected
dataset.

# Introduction

Humans with Sequence Space Synesthesia (SSS) represent ordered sequences
in particular spatial positions. For example, August (i.e. the
*inducer)* might be represented in the bottom left position (i.e. the
position is here the *concurrent),* this position is relative to the
concurrent position of the other months which could form a circle all
together. In addition time units, numbers also take particular forms
[@galton1880], not to be confused with the Mental Number Line \[Dehaene
to be found\]. These forms are idiosyncratic, meaning they might vary
across individuals. This makes it difficult to detect authentic SSS and
therefore give precise estimates of prevalence in the population.
Estimated prevalence for SSS in the general population spans between 4.4
% [@brang2013] and 14.2 % [@seron1992], see also [@ward2018;
@sagiv2006]. Hence a reliable diagnostic tool to detect SSS would also
be useful to investigate SSS.

Diagnostic depends on the definition of the conditions under
investigation. A strict definition of Synestheesia requires five
different criteria [@deroy2013]. *Automaticity*: the *inducer*
automatically triggers the *concurrent*. For example august might
automatically trigger it's specific spatial location.
*Unidirectionality*: while the *inducer* triggers the concurrent, the
concurrent does not trigger the inducer. Hence the bottom left position
doe not trigger August. *Consciousness*: The concurrent is consciously
percieved. *Developmentally early*: the experience was already present
during childhood. *Consistency*: the inducer-concurrent pair remains
stable in time. For example, August triggers the same bottom left
position. Consistency is arguably the most suited criteria to develop a
diagnostic tool since it is relatively simple to implement in a
behavioral task and quantify.

Hence given consistency, similar concurrent responses triggered by the
same inducers can be used as a marker for authentic SSS. Consistency
test have become golden standart for colour-grapheme synesthesia, where
an inducer is presented (i.e. letters of the alphabet) and the
participant is requested to selected the concurrent colour, using a
colour picker. Individual consistency is then calculated as the distance
between repeated colour responses to the same inducers. Interestingly,
the best colour space to detect colour-grapheme synesthesia is CIE\*LUV,
a colour space developed to be isoform to human perception
[@rothen2013]. Analogously to grapheme-colour synesthetes, consistency
test can be used to diagnose SSS. In that tasks, it is repeatedly asked
to report the position of the inducers on a screen. The total area
between the responses of same inducer (i.e. a triangle if repeated three
times) is then used as characteristic to diagnose SSS. The rationale
being that consistent responses would lead to smaller area than
inconsistent ones [@rothen2016]. This method resembles how number forms
are describe in the single case study [@piazza2006], see Experiment 1.

However characterizing synesthetes from non synesthetes using total area
has several limitations. For example high consistency by non-synesthetes
can be achieved by giving all responses on the same screen position
(i.e. false positive). Moreover, this kond of criteria might bias the
diagnosis to include synesthetes with straigth lines which leads to less
variability than more complex forms[@Ward].

The goal of the present registered report is to first identify new
features characterizing synesthetes responses based on already available
datasets and test the best working features on a future dataset. The new
features are desgined to take advantage of two properties of synesthetic
responses that have not been included in precedent consistency tests.
First, sequentiality on top of single inducer responses the ordered
position between subsequent induces is important. For example the
relative position of August and the other months. From numerical
cognition, ordinality has been aknowledged to be an important semantic
property of numbers, also given their sequential acquisition (i.e. 1 is
learned before 2). Second, thee particular synthetic forms of the
sequential spatial location. These forms might have geometrical
properties. For example months of the year might be represented
circularly (as already described by [@galton1880] for numbers).

To take advantage of sequential and geometrical synesthetic forms, we
harnessed a geo-spatial package[@pebesma2018] to extract geometrical
features from participant x and y coordinate responses. This packages
allows for example to build string or polygons for each repetition and
compare different geometrical features. Those individual geometrical
features are then compared using Receiver Operator Charactheristics
(ROC) between individuals grouped as synesthes and control. In the
present *phase I*,we compare ROC on three merged derivationdatasets
using the same task on SSS [@rothen2016, @ward; @vanpetersen2020a]. In
future *phase II*, we compare whether the features selected to diagnose
SSS in *phase I,* on a validation dataset that is not yet acquired
(registered report on the open science foundation:
<https://osf.io/9efjb/>**).**

# Methods

*Phase I: present analyses*. We merge three available dataset and
compared available diagnostic criteria across datasets using Reciever
Operator Charachteristics (ROC) for different approaches. A first
approach of diagnostic criteria based on a stimulus level consistency.
The rationale being that synesthetes should produce more consistent
(i.e. less variable) across repetitions. Such criteria include area
between repetitions[^1], standard deviation of responses \[\@ward\] or
permuted responses [@root2025]? . The second suggested approach is to
look at the geometrical form generated by the responses across
repetitions. The rationale here is that synestetic responses should have
geometrical feature that differ from controls. For example, several SSS
representations for months are circular. For the form

[^1]: Of which we replicate the original results [@rothen2016] in a
    separate document

Third, we compare wheter the featuers lead to somilar ROC
charachteristics across the different sets (i.e. for months, weeks and
numbers). Fourth, we compute new candidate geometrical features that
could be used to diagnose SS. Finally we summarize and compare all ROC
and select the best features that class synesthetes from control with
the merged dataset.

*Phase II: future analyses.* On a future dataset using the same task, we
will compare the predictive power of the selected features using ROC.

## Materials

A the exception of [@rothen2016] (see
<https://osf.io/6hq94/files/osfstorage>), the data from [@ward;
@vanpetersen2020a] were collected online. The 29 inducers were: the 12
months of a year, 7 days of the week and 10 numbers (i.e. hindo-arabic
numerals from 0 to 9). [@vanpetersen2020a] Also presented 50 and 100
numerals, which we excluded here. [@ward] data is collected using the
Syntoolkit.

## Procedure

The details for each procedure is described in each respective article
[@rothen2016; @ward; @vanpetersen2020a], here we describe the common
task.

Each participant is presented with one one inducer at a time at the
center of a otherwise white screen. The participant is instructed to
click at the screen position that they visualize them. Inducers order is
randomized and each inducer is repeated three times.

*The order of the stimuli was randomised, but such that no stimulus was
repeated until the previous batch of unique stimuli (N = 29) had been
presented.*

# *Phase I* Methods

The data for *phase I,* comes from: [@rothen2016],[@ward] (from:
[https://osf.io/p5xsd/files/osfstorage](#0){.uri}) and
[@vanpetersen2020]

-   Root [@root2021]

```{r Load data}
#| warning: false
# In the following I upload and merge the data from Ward, Rothen and Van Peters. Data is stored into a full dataset ds (i.e. 1 row per trial) and a dataset per participant ds_Quest (i.e. 1 row per participant).

### 0.1.1. Ward Data

ds_syn       <- read_excel("raw_synaesthetes_consistency_anon.xlsx")
ds_syn$group <- "Syn"
ds_ctl       <- read_excel("raw_controls_consistency_anon.xlsx")
ds_ctl$group <- "Ctl"

ds_Q_syn       <- read_excel("raw_synaesthetes_questionnaire_anon.xlsx")
ds_Q_syn$group <- "Syn"
ds_Q_ctl       <- read_excel("raw_controls_questionnaire_anon.xlsx")
ds_Q_ctl$group <- "Ctl"

# Merge wards datafiles:
ds <- merge(ds_syn,ds_ctl, all = TRUE)
ds_Q <- merge(ds_Q_syn,ds_Q_ctl, all = TRUE)

# Ward only uses those who completed the Questionnaire (i.e. N = 215+252 = 467)
ds <- ds %>% 
  filter(session_id %in% unique(ds_Q$session_id))
 
### 0.1.1. Rothen Data

ds_Rothen <- read.csv("~/Documents/SpaceSequenceSynDiagnostic/SpaceSequenceSynDiagnostic/rawdata.txt", sep="")

### 0.1.2. Rename variables to match datasets

sum(ds_Q$consistency_score != ds_Q$consistency) # Duplicate variable
ds_Q$consistency <- NULL
ds_Q$...36 <- NULL
ds_Q$...37 <- NULL
ds_Q$mean_simulation_Z <- NULL
ds_Q$SD_simulation <- NULL
ds_Q$`z-score`  <- NULL

rm(ds_syn,ds_ctl,ds_Q_syn,ds_Q_ctl)

ds$ID <- ds$session_id
ds_Q$ID <- ds_Q$session_id

ds_Q$dataSource <- "Ward"
ds$dataSource <- "Ward"

# From Rothen:
names(ds_Rothen)[names(ds_Rothen) == "Group"] <- "group"
ds_Rothen$group <- as.factor(ds_Rothen$group)
levels(ds_Rothen$group) <- c("Ctl","Syn")
names(ds_Rothen)[names(ds_Rothen) == "Inducer"] <- "stimulus"
names(ds_Rothen)[names(ds_Rothen) == "X"] <- "x"
names(ds_Rothen)[names(ds_Rothen) == "Y"] <- "y"
ds_Rothen$SynQuest <- ds_Rothen$group == "Syn"

ds_Rothen$dataSource <- "Rothen"

# From the paper (all the same since lab based):
ds_Rothen$width <- 1024
ds_Rothen$height <- 768
```

```{r, Merge1 Rothen and Ward data}
## 0.2 Merge data:
# remove non matching colnames: 

ColNames_ds <- colnames(ds_Rothen)[colnames(ds_Rothen) %in% colnames(ds)]

ds <- ds %>%
  select(all_of(ColNames_ds))
ds_Rothen <- ds_Rothen %>%
  select(all_of(ColNames_ds))

# Data:
ds   <- merge(ds,ds_Rothen, all = TRUE)

# Because there is no questionnaire in Nicola's data:
ds$SynQuest[ds$dataSource == "Rothen"] = "NaN"

# Questionnaire:
ID <- unique((ds_Rothen$ID))
ds_Q_Rothen <- as.data.frame(ID)
ds_Q_Rothen$dataSource <- "Rothen"
ds_Q_Rothen <- merge(ds_Q_Rothen, ds_Rothen %>% group_by(ID) %>% select(ID, dataSource, group) %>% filter(row_number() == 1), by = c("ID","dataSource"))

# Append rows:
ds_Q$ID <- as.character(ds_Q$ID)
ds_Q <-  bind_rows(ds_Q,ds_Q_Rothen)

# Clear up:
rm(ds_Q_Rothen, ds_Rothen)

## 0.3 Wrangle dataset

# Add Condition, i.e. stim type:
ds$Cond <- NaN
ds$Cond[ds$stimulus %in% c("1","2","3","4","5","6","7","8","9","0")] <- "number"
ds$Cond[ds$stimulus %in% c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")] <- "weekday"
ds$Cond[ds$stimulus %in% c("January", "February", "March", "April", "May","June","July","August","September","October","November","December")] <- "month"
```

```{r, Merge2}
### ADD Van Petersen Cortex data
library(readxl)

filedir <- "~/Documents/SpaceSequenceSynDiagnostic/VanPetersen/Cortex/di.dcc.DSC_2018.00019_653/Consistency test/Preprocessed data/"
fn <- paste0(filedir,dir(filedir))

ds_PeterCor <- read_excel(fn[1],
                                col_types = c("text", "numeric", "text", 
                                              "text", "numeric", "numeric", "numeric", 
                                              "numeric", "numeric", "numeric", 
                                              "numeric", "numeric", "text", "text", 
                                              "text", "numeric", "numeric", "numeric"))

for(i in 2:length(fn)){
  ds_i <- read_excel(fn[i],
                     col_types = c("text", "numeric", "text", 
                                   "text", "numeric", "numeric", "numeric", 
                                   "numeric", "numeric", "numeric", 
                                   "numeric", "numeric", "text", "text", 
                                   "text", "numeric", "numeric", "numeric"))
    ds_PeterCor <- merge(ds_PeterCor, ds_i, all = TRUE)
}

# Here information about group:
ds_PeterCor_Q <- read_excel("~/Documents/SpaceSequenceSynDiagnostic/VanPetersen/Cortex/di.dcc.DSC_2018.00019_653/Consistency test/Final data files/Consistency_scores.xlsx")

names(ds_PeterCor_Q)[names(ds_PeterCor_Q) == "PPcode"] <- "ID"
names(ds_PeterCor)[names(ds_PeterCor) == "Code"] <- "ID"

# Exclude and Match ID's:
# IDExcl <- unique(ds_PeterCor$ID[is.na(ds_PeterCor$MouseClick.RESPCursorX)])
# ds_PeterCor <- ds_PeterCor %>%
#     filter(!ID %in% IDExcl)

ID_IN <- intersect(ds_PeterCor$ID, ds_PeterCor_Q$ID)
ds_PeterCor <- ds_PeterCor %>%
    filter(ID %in% ID_IN)
ds_PeterCor_Q <- ds_PeterCor_Q %>%
    filter(ID %in% ID_IN)

# Merge data with group and all dataset: 
ds_PeterCor <- ds_PeterCor_Q %>%
  left_join(ds_PeterCor, by = "ID")

# Rename to match:
names(ds_PeterCor)[names(ds_PeterCor) == "Group"] <- "group"
ds_PeterCor[ds_PeterCor$group == "SSS",]$group <- "Syn"
ds_PeterCor[ds_PeterCor$group == "control",]$group <- "Ctl"
names(ds_PeterCor)[names(ds_PeterCor) == "MouseClick.RESPCursorY"] <- "y"
names(ds_PeterCor)[names(ds_PeterCor) == "MouseClick.RESPCursorX"] <- "x"
names(ds_PeterCor)[names(ds_PeterCor) == "word"] <- "stimulus" # Will need to translate to english!
names(ds_PeterCor)[names(ds_PeterCor) == "BlockList.Cycle"] <- "repetition"

ds_PeterCor$width  <- 1920 # "screen with display resolution set to 1920 x 1080, controlled by a Dell computer running Windows 7."
ds_PeterCor$height <- 1080
ds_PeterCor$dataSource <- "PeterCor"

ds_PeterCor <- ds_PeterCor %>%
  select(all_of(ColNames_ds))

# Translate, weekdays:
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "maandag"]   <- "Monday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "dinsdag"]   <- "Tuesday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "woensdag"]  <- "Wednesday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "donderdag"] <- "Thursday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "vrijdag"]   <- "Friday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "zaterdag"]  <- "Saturday"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "zondag"]    <- "Sunday"

# Translate, Monthc:
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "januari"]   <- "January"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "februari"]  <- "February"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "maart"]     <- "March"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "april"]     <- "April"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "mei"]       <- "May"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "juni"]      <- "June"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "juli"]      <- "July"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "augustus"]  <- "August"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "september"] <- "September"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "oktober"]   <- "October"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "november"]  <- "November"
ds_PeterCor$stimulus[ds_PeterCor$stimulus == "december"] <- "December"

# They added 2 numbers that are not in the other datasets:
ds_PeterCor <- ds_PeterCor %>%
    filter(stimulus != "50") %>%
    filter(stimulus != "100")

##### MERGE ####
ds   <- merge(ds,ds_PeterCor, all = TRUE)

# Questionnaire:
names(ds_PeterCor_Q)[names(ds_PeterCor_Q) == "Group"] <- "group"
ds_PeterCor_Q[ds_PeterCor_Q$group == "SSS",]$group <- "Syn"
ds_PeterCor_Q[ds_PeterCor_Q$group == "control",]$group <- "Ctl"

ds_PeterCor_Q$dataSource <- "PeterCor"

ds_PeterCor_Q <- ds_PeterCor_Q %>%
  select(ID, group, `Consistency All`, `Consistency Days`, `Consistency Months`,`Consistency Numbers`, dataSource)

# Append rows:
ds_Q <-  bind_rows(ds_Q,ds_PeterCor_Q)

# Tidy workspace:
rm(ds_PeterCor_Q,ds_PeterCor,ds_i, ID, fn,ColNames_ds,i,ID_IN)
```

```{r Wrangle data}
# Now we can enrich our dataset and process several checks
# Add Condition, i.e. stim type:
ds$Cond <- NaN
ds$Cond[ds$stimulus %in% c("1","2","3","4","5","6","7","8","9","0")] <- "number"
ds$Cond[ds$stimulus %in% c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")] <- "weekday"
ds$Cond[ds$stimulus %in% c("January", "February", "March", "April", "May","June","July","August","September","October","November","December")] <- "month"

# Add stimulus repetition number
ds <- ds %>%
  group_by(ID, stimulus) %>%
  arrange(ID, stimulus, .by_group = TRUE) %>%
  mutate(repetition = row_number()) %>%
  ungroup()

# Factorialize group variable to avoid confusions:
ds$group   <- as.factor(ds$group)
ds_Q$group <- as.factor(ds_Q$group)
```

```{r, Filter data}

## 0.3 Filter trials
n_trials1 <- length(ds$stimulus)
# Remove if not 3 repetitions per stimuli:
ds <- ds %>% 
    group_by(ID,Cond,stimulus) %>% 
    mutate(Nrep = length(stimulus))

ds <- ds %>%
  filter(Nrep == 3)

n_trials2 <- length(ds$stimulus)

# Sanity Check (should be empty)
# tmp <- ds %>% filter(repetition > 3)

# Compute mean x,y:
ds <- ds %>% 
  group_by(ID, Cond, stimulus) %>%
  mutate(X_mean = mean(x), Y_mean = mean(y)) 

# Sanity Check:
# ds %>% group_by(dataSource) %>% summarise(n = length(stimulus), maxrep = max(Nrep), minrep = min(Nrep))

# First we excluded all stimuli that were not repeated 3 times, going from `r ntrials1` to `r ntrials2` trials.

## 0.4 Filter ID's
# Match ID's across datasets:
ID_ds   <- unique(ds$ID)
ID_ds_Q <- unique(ds_Q$ID)

ds <- ds %>%
    filter(ID %in% ID_ds[ID_ds %in% ID_ds_Q]) %>%
    filter(ID %in% ID_ds_Q[ID_ds_Q %in% ID_ds])

ds_Q <- ds_Q %>% 
    filter(ID %in% ID_ds[ID_ds %in% ID_ds_Q]) %>%
    filter(ID %in% ID_ds_Q[ID_ds_Q %in% ID_ds])

ID_onlyQ <- ID_ds[! ID_ds %in% ID_ds_Q]

# Sanity Check:
# sum(ID_ds == ID_ds_Q) == length(unique(ds$ID))
# sum(ID_ds == ID_ds_Q) == length(unique(ds_Q$ID))

# Important. To reproduce Rothen, the X = -1 and Y = -1 need to be NaN:

ds$x[ds$x == -1] <- NaN
ds$y[ds$y == -1] <- NaN

# Eventually to (need to check)
# ds$x[ds$x == 0] <- NaN
# ds$y[ds$y == 0] <- NaN
# Why avoid this:
# If not able to reproduce the figure, the thing is that 0,0 coordinates might be genuine responses.
# ds %>%
#     filter(ID %in% 59854) %>%
#     select(nLineCross, stimulus, Cond, repetition, x_zs,y_zs,y,x) %>%
#     filter(Cond %in% "number") %>%
#     group_by(stimulus) %>%
#     arrange(stimulus) %>%
#     arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
#     arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
#     ungroup()  %>%
#     group_by(ID, Cond,repetition)  %>%
#     mutate(nLineCross2 = (count_self_intersections(x_zs,y_zs, verbose = FALSE))) %>%
#     ggplot(aes(x = x_zs, y = y_zs, group = stimulus, label = nLineCross2, fill = stimulus)) +
#     geom_path(aes(x = x_zs, y = y_zs, group = repetition), alpha = 0.2) +
#     geom_text(aes(x = 0, y = 2)) +
#     facet_grid( ~ repetition) +
#     theme_minimal()

# Detect ID that with overlapping coordinates across responses:
tmp <- ds %>%
    group_by(ID,Cond, repetition) %>%
    mutate(SameCoord_x = length(unique(x[!is.na(x)])) == 1,SameCoord_y = length(unique(x[!is.na(y)])) == 1)

ID_same_x <- tmp %>% 
  filter(SameCoord_x) %>%
  group_by(ID) %>%
  filter(row_number() == 1) %>%
  select(ID)

ID_same_y <- tmp %>% 
  filter(SameCoord_y) %>%
  group_by(ID) %>%
  filter(row_number() == 1) %>%
  select(ID)

Same_x_same_y <- ID_same_y$ID[ID_same_y$ID %in% ID_same_x$ID]

# Could it skew a test? Yes:
ds %>%
    filter(ID %in% Same_x_same_y) %>%
    group_by(ID) %>%
    filter(row_number() == 1) %>%
    ungroup() %>%
    count(group)

ds$isSame_xy   <- ds$ID %in% Same_x_same_y
ds_Q$isSame_xy <- ds_Q$ID %in% Same_x_same_y
# Data Example
# ID_sample <- sample(Same_x_same_y,1)
# ds %>%
#     filter(ID %in% ID_sample) %>%
#     select(x,y, Cond) %>%
#     print(n = 200)

# Visualize:
# ds %>%
#   filter(ID %in% Same_x_same_y) %>%
#   group_by(stimulus) %>%
#   arrange(stimulus) %>%
#   arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
#   arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
#   ggplot(aes(x = x, y = y, group = stimulus, fill = stimulus, colour = as.factor(repetition), label = stimulus)) +
#   geom_path(aes(x = x, y = y, group = repetition), alpha = 0.2) +
#   geom_text() +
#   facet_grid(Cond ~ ID) +
#   theme_minimal()

# Honestly, this one should not be in the data:
# ds %>% filter(ID == "12_CH")%>% select(Cond,x,y)  %>% print(n = 87) 
# Only responses for months.
```

```{r Add_zs}
# The main difficulty is that z-score need variability. Now if ID clicked on the same coordinates, no zs can be calculated. So what should it be? 0.

## 0.4 Standardize/scale coordinates
ds <- ds %>%
  group_by(ID) %>% # Not by Cond, so to avoid NaN's
  mutate(x_zs = scale(x)) %>%
  mutate(y_zs = scale(y))

# ATTENTION WHEN INTERPRETING THIS:
ds$x_zs[is.na(ds$x_zs)] <- 0
ds$y_zs[is.na(ds$y_zs)] <- 0
```

```{r, FilterID}

# The problem is that some (i.e. length(NaN_IDList)) ID have clicked always at the same coordinates in some conditions:
# NaN_IDList <- ds %>% 
#     filter(x_zs == "NaN") %>%
#     select(ID)
# 
# # 2 ID to exclude
# NaN_IDList <- unique(NaN_IDList$ID)

# I.e.:
# ds %>% 
#     group_by(Cond) %>%
#     filter(ID == sample(NaN_IDList,1)) %>%
#     filter(row_number() %in% 1:5)
# I can't replace them with 0 or averages, since it would bias the whole consistency computation! Should exclude them. 

# 2 ID's responded always in the same position, exclude them:
# ds <- ds %>%
#     filter(!ID %in% NaN_IDList)

# Mean zs across repetitions: (i.e. centroid between repetitions)
ds <- ds %>% 
  group_by(ID, Cond, stimulus) %>%
  mutate(X_mean_zs = mean(x_zs), Y_mean_zs = mean(y_zs))

# We exclude `r length(NaN_IDList)` participants for which we could not
# compute the *z-scores*.
```

```{r ManualAdjust}
# Manually adjust pixels:
# Note: 29 since 29 stimuli
ds$height[ds$ID == 29324] <- 1080
ds$width[ds$ID == 32190 ] <- 1440

# This ID has unexpected changing screen settings, consider exclusion:
ds$width[ds$ID == 33168 ]  <- 308
ds$height[ds$ID == 33168 ] <- 149

ds$width[ds$ID == 35556 ]  <- 1439
ds$height[ds$ID == 35556 ] <- 734

ds$width[ds$ID == 48114 ]  <- 1593
ds$height[ds$ID == 48114 ] <- 671

ds$width[ds$ID == 59854]  <- 1366
ds$height[ds$ID == 59854] <- 663

ds$width[ds$ID == 63127]  <- 1920
ds$height[ds$ID == 63127] <- 880

# Sanity Check: should be empty:
# ID_heigthKO <- ds %>%
#   group_by(ID) %>%
#   mutate(nhig = length(unique(height))) %>%
#   filter(nhig != 1) %>%
#   filter(row_number()==1) %>%
#   pull(ID)
# 
# ID_widthKO <- ds %>%
#   group_by(ID) %>%
#   mutate(nwid = length(unique(width))) %>%
#   filter(nwid != 1) %>%
#   filter(row_number()==1) %>%
#   pull(ID)

ds$Screen_area <- ds$width*ds$height
```

## *Phase I.* Population

```{r PopulationTable}
#| eval: false
#| warning: false
#| include: false
knitr::kable(ds_Q %>%
    group_by(group, dataSource) %>%
    summarize(N = length(ID)) %>%
    pivot_wider(names_from = group, values_from = N))
```

For all the three datasets, the synesthetes were self-reported.

| Source | Synesthetes | Controls | Total | Age | n females |
|------------|------------|------------|------------|------------|------------|
| [@rothen2016] | 33 | 37 | 70 | 23.1; 28.2 (ctl) | 24; 27 (ctl) |
| [@vanpetersen2020] | 23 | 21 | 44 | 23.22; 21.57 (ctl) | 20; 19 (ctl) |
| [@wardb] | 252 | 215 | 467 | 37.21; 19.90 (ctl) | 202; 178 (ctl) |
|  |  |  |  |  |  |
| Merged | 308 | 273 | 557 |  |  |

## *Phase I*. Analysis

First, we replicate consistency methods found in the literature using
the same task ([@rothen2016; @ward; @vanpetersen2020a; @root2021]) and
compare the results.

Second, we extract features based on the form. (C) We harness a
geography package to compute segment based features (D) We compute
polygon based features. (E) Convex Hull (F) Angles.

# *Phase I.* Results Reproduce

# Replicated features

## Standart deviation

```{r}
# Rescale x & y coordinates depending on screen size:
ds$xSc <- ds$x/ds$width
ds$ySc <- ds$y/ds$height

# Compute the SD across all trials (per ID):
ds <- ds %>% 
  ungroup() %>%
  group_by(ID) %>%
  mutate(SD_ID_xsc = sd(xSc, na.rm = TRUE)) %>%
  mutate(SD_ID_ysc = sd(ySc, na.rm = TRUE)) 

# Add to ds_Q:
ds_Q <- merge(ds_Q, ds %>% ungroup() %>% group_by(ID) %>% arrange(ID) %>% filter(row_number() == 1) %>% select(ID,SD_ID_xsc,SD_ID_ysc),by = "ID")

# Sanity Check 
# plot(ds_Q$x_sd, ds_Q$SD_ID_xsc)
# plot(ds_Q$y_sd, ds_Q$SD_ID_ysc)
# NonMatchID <- unique(ds_Q$ID[round(ds_Q$x_sd,1) != round(ds_Q$SD_ID_xsc,1)])
# ds_Q %>%
#   filter(ID %in% NonMatchID) %>%
#   select(ID, x_sd, SD_ID_xsc) # OK!
```

## Triangle area Consistency

**Definition**: Calculating consistency Each stimulus is represented by
three xy coordinates - (x1, y1), (x2, y2), (x3, y3) - from the three
repetitions. For each stimulus, the area of the triangle bounded by the
coordinates is calculated as follows:\
$Area = (x1y2 + x2y3 + x3y1 – x1y3 – x2y1 – x3y2) / 2$

```{r, triangleArea_fun}

# Define area calculation function
triangle_area <- function(x, y) {
  if(length(x) != 3 | length(y) != 3) return(NA)
  area <- abs(
    x[1]*y[2] + x[2]*y[3] + x[3]*y[1] -
    x[1]*y[3] - x[2]*y[1] - x[3]*y[2]
  ) / 2
  return(area)
}

```

```{r, areaConsistency}
## Compute triangle area by group:
ds <- ds %>%  
  group_by(ID, stimulus) %>%
  mutate(triangle_area = triangle_area(x, y)) %>%
  ungroup()

## Summarize By ID:
ds_ID <- ds %>%
  ungroup() %>% group_by(ID,group) %>%
  summarize(triangle_area_GA = mean(triangle_area, na.rm = TRUE))  %>%
  select(ID,triangle_area_GA)

# tmp_perID <- ds  %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, Consistency)
# names(tmp_perID) <- c("ID", "Consistency2")
ds_Q <- merge(ds_Q,ds_ID,by = "ID")
rm(tmp_perID)

```

```{r, area_zsConsitency}

ds <- ds %>%  
  group_by(ID, stimulus) %>%
  mutate(triangle_area_zs = triangle_area(x_zs, y_zs)) %>%
  ungroup()

ds <- ds %>%  
   group_by(ID,repetition) %>%
   mutate(Consistency_zs = (sum(triangle_area_zs)))

tmp_perID2 <- ds %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, Consistency_zs)

ds_Q <- merge(ds_Q,tmp_perID2,by = "ID")
rm(tmp_perID2)
```

# *Phase I.* Results: Novel features

## Segment self-intersection

The first feature is self-intersections. For each condition and
repetition separately, we count how many times the segments
self-intersects. Then we average across conditions and repetitions to
have one value per subject. Hence the value represents the average
number of self-intersections per segment a subject has.

```{r, selfInter_fun}

# Define function: this was generated by chatgpt and it does the correct job after checking

count_self_intersections <- function(x, y, verbose = TRUE) {
  n <- length(x)
  if (n < 4) {
    if (verbose) cat("Need at least 4 points to check for self-intersection.\n")
    return(0)
  }

  # Orientation function
  orientation <- function(p, q, r) {
    val <- (q[2] - p[2]) * (r[1] - q[1]) - (q[1] - p[1]) * (r[2] - q[2])
    if (is.na(val)) return(NA)
    if (val == 0) return(0)
    if (val > 0) return(1) else return(2)
  }

  # Check if q lies on segment pr
  on_segment <- function(p, q, r) {
    if (any(is.na(c(p, q, r)))) return(FALSE)
    q[1] <= max(p[1], r[1]) && q[1] >= min(p[1], r[1]) &&
      q[2] <= max(p[2], r[2]) && q[2] >= min(p[2], r[2])
  }

  # Main intersection check
  segments_intersect <- function(p1, p2, p3, p4) {
    o1 <- orientation(p1, p2, p3)
    o2 <- orientation(p1, p2, p4)
    o3 <- orientation(p3, p4, p1)
    o4 <- orientation(p3, p4, p2)

    if (any(is.na(c(o1, o2, o3, o4)))) return(FALSE)

    # General case
    if (o1 != o2 && o3 != o4) return(TRUE)

    # Special colinear cases
    if (o1 == 0 && on_segment(p1, p3, p2)) return(TRUE)
    if (o2 == 0 && on_segment(p1, p4, p2)) return(TRUE)
    if (o3 == 0 && on_segment(p3, p1, p4)) return(TRUE)
    if (o4 == 0 && on_segment(p3, p2, p4)) return(TRUE)

    return(FALSE)
  }

  count <- 0
  for (i in 1:(n - 2)) {
    for (j in (i + 2):(n - 1)) {
      if (j == i + 1) next  # skip adjacent segments

      p1 <- c(x[i], y[i])
      p2 <- c(x[i + 1], y[i + 1])
      p3 <- c(x[j], y[j])
      p4 <- c(x[j + 1], y[j + 1])

      if (segments_intersect(p1, p2, p3, p4)) {
        count <- count + 1
        if (verbose) {
          cat(sprintf("Intersection #%d: segments (%d-%d) and (%d-%d)\n", count, i, i+1, j, j+1))
        }
      }
    }
  }

  if (verbose) cat("Total crossings:", count, "\n")
  return(count)
}

```

```{r, Self_Intersections}
# Number of intersections for each ID X Cond X repetition
# Important! The dataset must be correctly informed about stimulus order.
ds <- ds %>% 
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ungroup() %>%
  group_by(ID, Cond,repetition) %>%
  mutate(nLineCross = (count_self_intersections(x,y, verbose = FALSE))) %>%
  arrange(ID)# could also be x_zs and y_zs
```

```{r}

# Average self-intersections per ID by Cond X repetition
tmp <- ds %>%
  group_by(ID,Cond,repetition) %>%
  filter(row_number() == 1) %>% # keep only 1 row per IDXCondXrep
  ungroup() %>%
  group_by(ID) %>%
  mutate(SelfInter_persegm = mean(nLineCross)) %>%
  filter(row_number() == 1) %>% # keep only 1 row per ID
  select(ID, SelfInter_persegm) 


ds <- merge(ds, tmp,by = "ID")


ds_Q <- right_join(ds_Q, 
           (ds %>% 
                ungroup() %>% 
                group_by(ID) %>% 
                filter(row_number() == 1) %>%
                select(ID, SelfInter_persegm)), by = "ID")
```

## Segments (with sf)

We will take advantage of the `sf` package and connect the x and y
coordinates of ordered inducer with a segment. Sf hates NaN's. Either
convert them to 0 (as originally) or remove them. I'll start converting
to 0.

```{r}
library(sf)
# Turn off spherical geometry:
sf::sf_use_s2(FALSE)

# Explicitly enforce item order (i.e. ordinality):

ds_segm <- ds %>%
  group_by(ID, stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ungroup() %>%
  arrange(ID) %>%
  filter(!is.nan(x_zs), !is.nan(y_zs)) %>% # sf hates NaN! Not needed anymore, NaN's are managed above in the code.
  mutate(
    group = as.character(group),
    ID     = as.character(ID),
    Cond      = as.character(Cond),
    repetition     = as.integer(repetition),
    dataSource = as.character(dataSource)
  ) %>%
  # arrange(ID, Cond, repetition,group) %>%
  group_by(ID, Cond, repetition,group) %>%
  summarise(
    geometry = st_sfc(st_linestring(as.matrix(cbind(x_zs, y_zs)))), # preserves order
    .groups = "drop"
  ) %>%
  st_as_sf(crs = NA)
```

### Segment length (should replicate Rothen)

```{r}
ds_segm$leng <- st_length(ds_segm)

ds_segm <- ds_segm %>%
  group_by(ID) %>%
  mutate(GA_segm_leng = mean(leng, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                ds_segm %>% 
                ungroup() %>% 
                group_by(ID) %>% 
                filter(row_number() == 1) %>% 
                st_drop_geometry() %>%
                select(ID, GA_segm_leng),by = "ID")

```

# Distances between repetitions

```{r}
ID_list     <- unique(ds$ID)
Cond_list   <- unique(ds$Cond)
ds$BtwDist <- NaN

for (ID_n in 1:length(ID_list)) {
   for (Cond_n in 1: length(Cond_list)) {
     ds_here = ds_segm %>% filter(ID == ID_list[ID_n]) %>% filter(Cond == Cond_list[Cond_n])
     if (length(ds_here$geometry) != 3) {
       break
     }

     ds[ds$ID == ID_list[ID_n] & ds$Cond == Cond_list[Cond_n],]$BtwDist = mean(st_distance(ds_here))
    
   }
}


# Average per ID
ds <- ds %>%
  group_by(ID) %>%
  mutate(GA_BtwDist = mean(BtwDist, na.rm = TRUE))


ds_Q <- right_join(ds_Q, 
                   ds %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, GA_BtwDist),
                   by = "ID")
```

## Polygon based geometries

### Polygon area

```{r}
ds_poly          <- st_cast(ds_segm, "POLYGON")
ds_poly$area     <- st_area(ds_poly)

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(GA_areaPoly = mean(area, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, GA_areaPoly),
                   by = "ID")
```

# Polygon simplicity

```{r}
# Might depend on cast:
ds_poly          <- st_cast(ds_segm, "POLYGON", group_or_split = TRUE)

ds_poly$isSimple <- st_is_simple(ds_poly)
# st_is_simple returns a logical vector, indicating for each geometry whether it is simple (e.g., not self-intersecting)

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(GA_isSimple = mean(isSimple, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, GA_isSimple),
                   by = "ID")
```

# Topological validity Structure

is topologically valid:

From the package description: *"For projected geometries, st_make_valid
uses the lwgeom_makevalid method also used by the PostGIS command
ST_makevalid if the GEOS version linked to is smaller than 3.8.0, and
otherwise the version shipped in GEOS; for geometries having ellipsoidal
coordinates s2::s2_rebuild is being used."* From
<https://postgis.net/docs/ST_IsValid.html>: *value is well-formed and
valid in 2D according to the OGC rules.* (Open Geopsatial Consotrtium)

```{r}
ds_poly$isValidStruct <- st_is_valid(ds_poly, geos_method = "valid_structure")

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(Sum_isValidStruct = sum(isValidStruct, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, Sum_isValidStruct),
                   by = "ID")

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(GA_isValidStruct = mean(isValidStruct, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, GA_isValidStruct),
                   by = "ID")


```

## is clockwise

```{r}
ds_poly$isClockwise <- lwgeom::st_is_polygon_cw(ds_poly)

ds_poly <- ds_poly %>%
  group_by(ID) %>%
  mutate(Sum_isClockwise= sum(isClockwise, na.rm = TRUE))

ds_Q <- right_join(ds_Q, 
                   ds_poly %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, Sum_isClockwise),
                   by = "ID")
```

## Convex Hull Area

```{r}
ds_conv_hull <- ds_segm %>%
  mutate(geometry = st_convex_hull(geometry))
ds_conv_hull$area_convhull <- st_area(ds_conv_hull)

ds_conv_hull <- ds_conv_hull %>%
  group_by(ID) %>%
  mutate(GA_area_convhull = mean(area_convhull, na.rm = TRUE))


ds_Q <- right_join(ds_Q, 
                   ds_conv_hull %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% filter(row_number() == 1) %>% select(ID, GA_area_convhull),
                   by = "ID")
```

# Permutations

Until now the form based features are computed by chronologically
ordered repetitions. For example, Monday is repeated three times per ID.
The coordinates for Monday presented the first time will always be used
to form the segment/polygon with the Tuesday presented the first time.
However, for consistency, this should be independent from chronological
order. To circumvent this, we can permute the repetitions per
conditions. I predict the permuted averages of the same features should
give rise to better AUC. Since these are time consuming I will only
apply permutations on the criteria that are at stake:\
topological validity.

```{r}
#| eval: false
#| include: false

# Takes about 40 minutes on my machine. So pre-saved the output. 
set.seed(42)


ds <- ds %>%
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>%
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ungroup() %>%
  group_by(ID, group, Cond, repetition) %>%
  mutate(StimOrder = 1:length(stimulus))



# First define output matrix:

n_perms <- 100
perm_names <- paste0("n_perm", sprintf("%03d", 1:n_perms))  # X001 … X100

ds_perm <-  as.data.frame(
  matrix(NA_real_, nrow = length(unique(ds$ID)), ncol = n_perms,
         dimnames = list(unique(ds$ID), perm_names))
)

ID_list <- unique(ds$ID)
total = length(ID_list)*n_perms
pb <- txtProgressBar(min = 0, max = total, style = 3)
k <- 0


for(Perm_n in 1:n_perms){
  perm_here <- perm_names[Perm_n]
  for(ID_n in 1:length(unique(ds$ID))){
    
     
    ################ Extract data per ID: ################
    ds_ID <- ds %>%
      filter(ID %in% ID_list[ID_n]) %>%
      select(ID, group, stimulus, Cond, nLineCross, repetition, StimOrder,x_zs,y_zs,SelfInter_persegm)
    
     ################ Apply perumtation (sample) across repetitions################
     ds_ID <- ds_ID %>%
      group_by(stimulus) %>%
      mutate(repetition_perm = sample(repetition)) 
    
    ################ Pass to sf ################
    ds_ID_segm <- ds_ID %>%
      group_by(ID, stimulus) %>%
      arrange(stimulus) %>%
      arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>% 
      arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
      filter(!is.nan(x_zs), !is.nan(y_zs)) %>% # sf hates NaN! Not needed anymore, NaN's are managed above in the code.
      ungroup() %>%
      arrange(ID) %>%
      mutate(
        group = as.character(group),
        ID     = as.character(ID),
        Cond      = as.character(Cond),
        repetition_perm     = as.integer(repetition_perm)
      ) %>%
      group_by(ID, Cond, repetition_perm,group) %>%
      summarise(
        geometry = st_sfc(st_linestring(as.matrix(cbind(x_zs, y_zs)))), # preserves order
        .groups = "drop"
      ) %>%
      st_as_sf(crs = NA) 
    
    ################ Convert to poly and compute validity ################
    ds_ID_poly               <- st_cast(ds_ID_segm, "POLYGON")
    ds_ID_poly$isValidStructPerm <- st_is_valid(ds_ID_poly, geos_method = "valid_structure")
    
    ds_ID_poly <- ds_ID_poly  %>%
      mutate(isValidStructPerm = mean(isValidStructPerm, na.rm = TRUE)) %>%
      filter(row_number() == 1)
    
    ################ save the perumted results ################
        if(!rownames(ds_perm[ID_n,]) == ID_list[ID_n]){
      warning("ID names do not match")
    }
    
    ds_perm[ID_n,Perm_n] <- ds_ID_poly$isValidStructPerm
    k <- k + 1
    setTxtProgressBar(pb, k)
  }
}

# Now average the permutations: 

ds_perm$ID <-rownames(ds_perm)
ds_perm$group <- ds_perm$ID  %in% ds_Q$ID[ds_Q$group == "Ctl"] # is true if Ctl
ds_perm$group[ds_perm$group]  <- "Ctl"
ds_perm$group[ds_perm$group == "FALSE"] <- "Syn"

write.csv2(ds_perm, "permuted_isValid.csv")

```

```{r}
ds_perm <- read.csv2("permuted_isValid.csv")
# read.csv2("permuted_isValid.csv")
ds_Perm_perID <-  ds_perm %>% 
  rowwise() %>% 
  mutate(isValid_M_perm_ID = mean(c_across(n_perm001:n_perm100)), isValid_Med_perm_ID = median(c_across(n_perm001:n_perm100))) %>% 
  select (ID, group, isValid_M_perm_ID, isValid_Med_perm_ID)


ds_Q <- right_join(ds_Q, 
                   ds_Perm_perID %>% st_drop_geometry() %>% ungroup() %>% group_by(ID) %>% arrange(ID) %>% filter(row_number() == 1) %>% select(ID, isValid_M_perm_ID, isValid_Med_perm_ID),
                   by = "ID")
```

```         
```

# Compare all features:

## Compute all ROC

```{r, InitializeROC_fun}
## Define function to compute ROC:
Comp_ROC <- function(data, group_col, feature, ID, Ord){
  # Ord: must be: moreSyn or moreCtl. Wheter feature value is moer in Syn or ctl.
  if(!setequal(levels(data[[group_col]]), c("Syn","Ctl"))){
    return(warning("group order must be Syn Ctl"))
    break
  }
  
  if (Ord == "moreSyn") {
    Dirhere <- "<"
  } else if (Ord == "moreCtl") {
    Dirhere <- ">"
  } else {
    return(warning("Ord must be moreSyn or moreCtl"))  
  }
  
  ################ ROC analyses ################ 
  
  ROC_here <- pROC::roc(data[[group_col]] ~ data[[feature]], data, 
                  direction= Dirhere,
                  percent=TRUE,
                  # arguments for ci
                  ci=TRUE, boot.n=100, ci.alpha=0.9, stratified=FALSE,
                  # arguments for plot
                  plot=FALSE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
                  print.auc=TRUE, show.thres=TRUE, print.thres = "best", print.thres.best.method="youden")
  
  # Best threshold using Youden's J
  best_coords <- pROC::coords(ROC_here, "best", 
                              ret = c("accuracy","threshold", "sensitivity","specificity","ppv","npv"), 
                              best.method = "youden")
  

  auc_val <- as.numeric(pROC::auc(ROC_here))
  ci_auc  <- ci.auc(ROC_here)
  
  new_row <- data.frame(
    Feature   = feature,
    AUC        = round(auc_val, 4),
    threshold  = as.numeric(best_coords[["threshold"]]),
    sensitivity= as.numeric(best_coords[["sensitivity"]]),
    specificity= as.numeric(best_coords[["specificity"]]),
    ppv        = as.numeric(best_coords[["ppv"]]),
    npv        = as.numeric(best_coords[["npv"]]),
    ci_low     = as.numeric(ci_auc[1]),
    ci_high    = as.numeric(ci_auc[3]),
    stringsAsFactors = FALSE
  )

  ################ Contingency table ################ 
  
  if (Ord == "moreSyn") {
    data$diagnosis <- ifelse(data[[feature]] >= best_coords$threshold,  "Test : Syn", "Test : Ctl")
  } else if (Ord == "moreCtl") {
    data$diagnosis <- ifelse(data[[feature]] >= best_coords$threshold,  "Test : Ctl","Test : Syn")
  } else {
    warning("Ord must be moreSyn or moreCtl")  
  }
  
  tab_counts <- table(data[[group_col]], data$diagnosis)
  
  tab_percent <- prop.table(tab_counts, margin = 1) * 100
  
  result <- matrix(
    paste0(tab_counts, " (", round(tab_percent, 1), "%)"),
    nrow = nrow(tab_counts),
    dimnames = dimnames(tab_counts)
  )
  ################ Mosaic plot of the contigency table #############
  
  # gp1 <- data %>%
  #   # mutate(test_out = SD_ID_ysc > best_coords[["threshold"]]) %>%
  #   ggplot() +
  #   geom_mosaic(aes(x = product(diagnosis, [[group_col]]), fill = diagnosis)) +
  #   scale_fill_manual(values = c("#66C2A5","#FC8D62")) +
  #   geom_mosaic_text(aes(x = product(diagnosis, [[group_col]]), label = after_stat(.wt))) +
  #   theme_minimal()
  
  ################ General description ################ 
  # Attention, na removed  
  Descr_table <- data %>%
    group_by(!!sym(group_col)) %>%
    summarize(n = length(unique(!!sym(ID))), Mean = mean(!!sym(feature), na.rm = TRUE), SD = sd(!!sym(feature), na.rm = TRUE))

  ################ Return outputs ################ 
  return(list(ROC_properties = new_row, Coningency_table =result, Descr_table = Descr_table, ROC_curve = ROC_here))

}
```

```{r}
all_cols <- colnames(ds_Q)
feature_list      <- all_cols[43:57]
feature_direction <- c(rep("moreSyn",2),rep("moreCtl",5),rep("moreSyn",8))
# Visual check if correct matches:
rbind(feature_list,feature_direction)


# Now we will loop through features:
# INITIALIZE ROC_curves
ROC_curves <- list(Comp_ROC(ds_Q, "group", feature_list[1],"ID",feature_direction[1])$ROC_curve,Comp_ROC(ds_Q, "group", feature_list[1],"ID",feature_direction[1])$ROC_properties$Feature)

ROC_contingency <- list(Comp_ROC(ds_Q, "group", feature_list[1],"ID",feature_direction[1])$Coningency_table,Comp_ROC(ds_Q, "group", feature_list[1],"ID",feature_direction[1])$ROC_properties$Feature)

# Initialize dataframe to collect relevant ROC values:
All_ROC <- data.frame(Feature=character(),
                      AUC = character(),
                 threshold=integer(),
                 sensitivity=integer(),
                 specificity=integer(),
                 ppv = integer(),
                 npv = integer(),
                 high_ci = integer(),
                 low_ci = integer(),
                 stringsAsFactors=FALSE)

# Loop by features:
for(fit_n in 2:length(feature_list)){
  ROC_here <- Comp_ROC(ds_Q, "group", feature_list[fit_n],"ID",feature_direction[fit_n])
  
  ROC_curves      <- c(ROC_curves,list(ROC_here$ROC_curve,ROC_here$ROC_properties$Feature))
  All_ROC         <- rbind(All_ROC, ROC_here$ROC_properties)
  ROC_contingency <- c(ROC_contingency,list(ROC_here$Coningency_table,ROC_here$ROC_properties$Feature))
}

```

# Summaries

## Summary table:

```{r}
library(kableExtra)

All_ROC_round <- All_ROC %>% 
    mutate_if(is.numeric, round,2)
All_ROC_round <- All_ROC_round[order(All_ROC_round$AUC, decreasing = TRUE), ]


# Add colours for a nice table:
n_color <- length(All_ROC_round$Feature)

All_ROC_round[2] <- lapply(All_ROC_round[2], function(x) {
    cell_spec(x, bold = T, 
              background  = spec_color(x, end = 0.9,  palette = paletteer::paletteer_c("ggthemes::Red-Green Diverging", n_color), direction = -1 ),
    )
})

All_ROC_round[4] <- lapply(All_ROC_round[4], function(x) {
    cell_spec(x, bold = T, 
              background  = spec_color(x, end = 0.9,  palette = paletteer::paletteer_c("ggthemes::Red-Green Diverging", n_color), direction = -1 ),
    )
})

All_ROC_round[5] <- lapply(All_ROC_round[5], function(x) {
    cell_spec(x, bold = T, 
              background = spec_color(x, end = 0.9,  palette = paletteer::paletteer_c("ggthemes::Red-Green Diverging", n_color), direction = -1 ),
    )
})

kbl(All_ROC_round, escape = F, align = "c") %>%
  kable_classic("striped", full_width = F)
```

## Summary plot

```{r}

library(RColorBrewer)

ggroc(list(ROC_curves[[1]],ROC_curves[[3]],ROC_curves[[5]],ROC_curves[[7]],
           ROC_curves[[9]],ROC_curves[[11]],ROC_curves[[13]],ROC_curves[[15]],
           ROC_curves[[17]],ROC_curves[[19]],ROC_curves[[21]],ROC_curves[[23]],
           ROC_curves[[25]],ROC_curves[[27]],ROC_curves[[29]])) +
  geom_segment(aes(x = 0, xend = 100, y = 100, yend = 0),
               color="grey", size = 0.01) +
  scale_color_manual(labels = c(ROC_curves[[2]],ROC_curves[[4]],ROC_curves[[6]],
           ROC_curves[[8]],ROC_curves[[10]], ROC_curves[[12]],ROC_curves[[14]],
           ROC_curves[[16]],ROC_curves[[18]],ROC_curves[[20]],ROC_curves[[22]],
           ROC_curves[[24]],ROC_curves[[26]],ROC_curves[[28]],ROC_curves[[30]]), 
           values =pals::cols25(15)) + 
  theme_minimal()

```

```{r}
#| eval: false
#| include: false
# Summary contingency:
Tab_index <- seq(from = 1, to = length(feature_list)*2-1, by = 2)
Feat_index <- seq(from = 2, to = length(feature_list)*2, by = 2)

for(i in 1:length(Feat_index)){
  print(ROC_contingency[[Feat_index[i]]])
  print(ROC_contingency[[Tab_index[i]]])
}

```

# Summary excluding overlapping responses

```{r}

ds_Q2 <- ds_Q %>% filter(!isSame_xy)

# Now we will loop through features:
# INITIALIZE ROC_curves
ROC_curves2 <- list(Comp_ROC(ds_Q2, "group", feature_list[1],"ID",feature_direction[1])$ROC_curve,Comp_ROC(ds_Q2, "group", feature_list[1],"ID",feature_direction[1])$ROC_properties$Feature)

ROC_contingency2 <- list(Comp_ROC(ds_Q2, "group", feature_list[1],"ID",feature_direction[1])$Coningency_table,Comp_ROC(ds_Q2, "group", feature_list[1],"ID",feature_direction[1])$ROC_properties$Feature)

# Initialize dataframe to collect relevant ROC values:
All_ROC2 <- data.frame(Feature=character(),
                      AUC = character(),
                 threshold=integer(),
                 sensitivity=integer(),
                 specificity=integer(),
                 ppv = integer(),
                 npv = integer(),
                 high_ci = integer(),
                 low_ci = integer(),
                 stringsAsFactors=FALSE)

# Loop by features:
for(fit_n in 2:length(feature_list)){
  ROC_here2 <- Comp_ROC(ds_Q2, "group", feature_list[fit_n],"ID",feature_direction[fit_n])
  
  ROC_curves2      <- c(ROC_curves2,list(ROC_here2$ROC_curve,ROC_here2$ROC_properties$Feature))
  All_ROC2         <- rbind(All_ROC2, ROC_here2$ROC_properties)
  ROC_contingency2 <- c(ROC_contingency,list(ROC_here2$Coningency_table,ROC_here2$ROC_properties$Feature))
}

```

```{r}

library(kableExtra)

All_ROC_round2 <- All_ROC2 %>% 
    mutate_if(is.numeric, round,2)
All_ROC_round2 <- All_ROC_round2[order(All_ROC_round2$AUC, decreasing = TRUE), ]


# Add colours for a nice table:
n_color <- length(All_ROC_round2$Feature)

All_ROC_round2[2] <- lapply(All_ROC_round2[2], function(x) {
    cell_spec(x, bold = T, 
              background  = spec_color(x, end = 0.9,  palette = paletteer::paletteer_c("ggthemes::Red-Green Diverging", n_color), direction = -1 ),
    )
})

All_ROC_round2[4] <- lapply(All_ROC_round2[4], function(x) {
    cell_spec(x, bold = T, 
              background  = spec_color(x, end = 0.9,  palette = paletteer::paletteer_c("ggthemes::Red-Green Diverging", n_color), direction = -1 ),
    )
})

All_ROC_round2[5] <- lapply(All_ROC_round2[5], function(x) {
    cell_spec(x, bold = T, 
              background = spec_color(x, end = 0.9,  palette = paletteer::paletteer_c("ggthemes::Red-Green Diverging", n_color), direction = -1 ),
    )
})

kbl(All_ROC_round2, escape = F, align = "c") %>%
  kable_classic("striped", full_width = F)
```

## Summary plot

```{r}

library(RColorBrewer)

# ggroc(list(ROC_curves2[[1]],ROC_curves2[[3]],ROC_curves2[[5]],ROC_curves2[[7]],ROC_curves2[[9]],ROC_curves2[[11]],
#            ROC_curves2[[13]],ROC_curves2[[15]],ROC_curves2[[17]],ROC_curves2[[19]],ROC_curves2[[21]])) +
#   geom_segment(aes(x = 0, xend = 100, y = 100, yend = 0),
#                color="grey", size = 0.01) +
#   scale_color_manual(labels = c(ROC_curves2[[2]],ROC_curves2[[4]],ROC_curves2[[6]],ROC_curves2[[8]],ROC_curves2[[10]],
#                               ROC_curves2[[12]],ROC_curves2[[14]],ROC_curves2[[16]],ROC_curves2[[18]],ROC_curves2[[20]],
#                               ROC_curves2[[22]]), values = brewer.pal(11,name = "Set3")) + 
#   theme_minimal()


ggroc(list(ROC_curves2[[1]],ROC_curves2[[3]],ROC_curves2[[5]],ROC_curves2[[7]],
           ROC_curves2[[9]],ROC_curves2[[11]],ROC_curves2[[13]],ROC_curves2[[15]],
           ROC_curves2[[17]],ROC_curves2[[19]],ROC_curves2[[21]],ROC_curves2[[23]],
           ROC_curves2[[25]],ROC_curves2[[27]],ROC_curves2[[29]])) +
  geom_segment(aes(x = 0, xend = 100, y = 100, yend = 0),
               color="grey", size = 0.01) +
  scale_color_manual(labels = c(ROC_curves2[[2]],ROC_curves2[[4]],ROC_curves2[[6]],
           ROC_curves2[[8]],ROC_curves2[[10]], ROC_curves2[[12]],ROC_curves2[[14]],
           ROC_curves2[[16]],ROC_curves2[[18]],ROC_curves2[[20]],ROC_curves2[[22]],
           ROC_curves2[[24]],ROC_curves2[[26]],ROC_curves2[[28]],ROC_curves2[[30]]), 
           values =pals::cols25(15)) + 
  theme_minimal()
```

# Combination ROC

```{r}


Model3 <- glm(group ~ isValid_M_perm_ID + Consistency_zs , data = ds_Q, family = binomial)
pred3 = predict(Model3)
ds_Q$Comb <- pred3
roc(ds_Q$group, ds_Q$Comb)

Comp_ROC(ds_Q, "group", "Comb","ID","moreSyn")$ROC_properties
ggroc(Comp_ROC(ds_Q, "group", "Comb","ID","moreSyn")$ROC_curve)

ds_Q2 <- ds_Q %>% filter(!isSame_xy)

Model3 <- glm(group ~ isValid_M_perm_ID + Consistency_zs , data = ds_Q2, family = binomial)
pred3 = predict(Model3)
ds_Q2$Comb <- pred3
roc(ds_Q2$group, ds_Q2$Comb)

Comp_ROC(ds_Q2, "group", "Comb","ID","moreSyn")$ROC_properties
ggroc(Comp_ROC(ds_Q2, "group", "Comb","ID","moreSyn")$ROC_curve)
```

# Could compare each feature singularly:

```{r}
ROC_Valid <- plot.roc(group ~ isValid_M_perm_ID , ds_Q, percent = TRUE, col="red")
ROC_Cons  <- lines.roc(group ~ Consistency_zs , ds_Q, percent = TRUE,col ="green")

roc.test(ROC_Valid, ROC_Cons)

ROC_Valid <- plot.roc(group ~ isValid_M_perm_ID , ds_Q2, percent = TRUE, col="red")
ROC_Cons  <- lines.roc(group ~ Consistency_zs , ds_Q2, percent = TRUE,col ="green")

roc.test(ROC_Valid, ROC_Cons)
```

## *Phase II* Methods

## *Phase II* Materials:

Materials are desribed here
<https://osf.io/pjb6e/?view_only=d467ebf4c1f94076ae4ac61298255065>.

## *Phase II* Planned population

[**https://osf.io/6h8dx**](https://osf.io/6h8dx)

# Discussion

From the different features we extracted, topological validity across
the repetitions appeared to be the one leading to the largest Area Under
the Curve. The optimal cutoff was exactly 1.5, leading to a sensitivity
() and specificity ().

The optimal criterion ineeds to be informed about the order between
inducers (i.e. to construct the polygons) and interestingly suggests
that synthetic inducer are structurally mapped following topological
rules analogous to geographical space structures. Hence suggesting a
spatial nature for the synthetic forms of space sequence synesthetes.

# Supplementary: Permute repetitions

Until now the form based features are computed by chronologically
ordered repetitions. For example, Monday is repeated three times per ID.
The cooridinates for monday presented the first time will alawys be used
to form the segment/polygon with the tuesday presented the first time.
However, for consistency, this should be independent from chronological
order. To circumvent this, we can permute the repetitions per
conditions. I predict the permuted averages of the same features should
give rise to better AUC.

```{r}
#| eval: false
#| include: false

ds <- ds %>%
  group_by(stimulus) %>%
  arrange(stimulus) %>%
  arrange(ordered(stimulus, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"))) %>%
  arrange(ordered(stimulus, levels = c("January", "February", "March", "April", "May","June","July","August","September","October","November","December"))) %>%
  ungroup() %>%
  group_by(ID, Cond, repetition) %>%
  mutate(StimOrder = 1:length(stimulus))



# First define output matrix:

n_perms <- 100
perm_names <- paste0("n_perm", sprintf("%03d", 1:n_perms))  # X001 … X100

ds_perm <-  as.data.frame(
  matrix(NA_real_, nrow = length(unique(ds$ID)), ncol = n_perms,
         dimnames = list(unique(ds$ID), perm_names))
)

ID_list <- unique(ds$ID)
total = length(ID_list)*n_perms
pb <- txtProgressBar(min = 0, max = total, style = 3)
k <- 0


for(Perm_n in 1:n_perms){
  perm_here <- perm_names[Perm_n]
  for(ID_n in 1:length(unique(ds$ID))){
    
    ds_ID <- ds %>%
      filter(ID %in% ID_list[ID_n]) %>%
      select(ID, stimulus, Cond, nLineCross, repetition, StimOrder,x,y,SelfInter_persegm)
    
    ds_ID <- ds_ID %>%
      group_by(stimulus) %>%
      mutate(repetition_perm = sample(repetition)) %>%
      group_by(repetition_perm,Cond) %>%
      mutate(nLineCross_perm = (count_self_intersections(x,y, verbose = FALSE)))
    
    if(!rownames(ds_perm[ID_n,]) == ID_list[ID_n]){
      warning("ID names do not match")
    }
    
    Out <- ds_ID %>%
      group_by(ID,Cond,repetition_perm) %>%
      filter(row_number() == 1) %>%
      ungroup() %>%
      group_by(ID) %>%
      mutate(SelfInter_persegm_perm = mean(nLineCross_perm)) %>%
      filter(row_number() == 1) %>%
      select(ID, SelfInter_persegm_perm) 
    # ds_ID$SelfInter_persegm_perm <- Out$SelfInter_persegm_perm
    
    ds_perm[ID_n,Perm_n] <- Out$SelfInter_persegm_perm
    k <- k + 1
    setTxtProgressBar(pb, k)
  }
}

# Now average the permutations: 

ds_perm$group <- rownames(ds_perm) %in% ds_Q$ID[ds_Q$group == "Ctl"] # is true if Ctl
ds_perm$group[ds_perm$group]  <- "Ctl"
ds_perm$group[ds_perm$group == "FALSE"] <- "Syn"

ds_Perm_perID <-  ds_perm %>% 
  rowwise() %>% 
  mutate(M_perm_ID = mean(c_across(n_perm001:n_perm100)), Med_perm_ID = median(c_across(n_perm001:n_perm100))) %>% 
  select (group, M_perm_ID, Med_perm_ID)

ds_Perm_perID$group <- as.factor(ds_Perm_perID$group)
Comp_ROC(ds_Perm_perID, "group", "M_perm_ID","ID","moreCtl")

ggroc(list(ROC_curves[[11]], Comp_ROC(ds_Perm_perID, "group", "M_perm_ID","ID","moreCtl")$ROC_curve))
All_ROC2[4,]
Comp_ROC(ds_Perm_perID, "group", "M_perm_ID","ID","moreCtl")$ROC_properties

# As predicted, the permutations lead to better criteria!
```

# Supplementary: Conditions

```{r}
All_ROC_Cond <- data.frame(Feature=character(),
                      AUC = character(),
                 threshold=integer(),
                 sensitivity=integer(),
                 specificity=integer(),
                 ppv = integer(),
                 npv = integer(),
                 high_ci = integer(),
                 low_ci = integer(),
                 stringsAsFactors=FALSE)
```

## Per Conditions

```{r, PerCondition}
ds <- ds %>%  
    group_by(ID,repetition,Cond) %>%
    mutate(Consistency_zs_Cond = (sum(triangle_area_zs)))

Cond_list <- unique(ds$Cond)
rm(ROC_curvesCond) # we are concatenating list, so make sure they don't already exist

for(Cond_i in 1:3) {
  PerCond <- ds %>%
    filter(Cond %in% Cond_list[Cond_i]) %>%
    Comp_ROC(., "group", "Consistency_zs_Cond","ID","moreCtl")
  
  PerCond$ROC_properties$Feature <- paste0(PerCond$ROC_properties$Feature, "_", Cond_list[Cond_i])
  All_ROC_Cond <- rbind(All_ROC_Cond, PerCond$ROC_properties)
  
  if(Cond_i == 1){
    # INITIALIZE ROC_curves
    ROC_curvesCond <- list(PerCond$ROC_curve,PerCond$ROC_properties$Feature)
  }
  
  ROC_curvesCond <- c(ROC_curvesCond,list(PerCond$ROC_curve,PerCond$ROC_properties$Feature))
}


ds_poly <- ds_poly %>%
  group_by(ID,Cond) %>%
  mutate(GA_isValidStructCond = mean(isValidStruct, na.rm = TRUE))

for(Cond_i in 1:3) {
  PerCond <- ds_poly %>%
    filter(Cond %in% Cond_list[Cond_i]) %>%
    mutate(group = as.factor(group)) %>%
    st_drop_geometry() %>%
    Comp_ROC(., "group", "GA_isValidStructCond","ID","moreCtl")
  PerCond$ROC_properties$Feature <- paste0(PerCond$ROC_properties$Feature, "_", Cond_list[Cond_i])
  All_ROC_Cond <- rbind(All_ROC_Cond, PerCond$ROC_properties)
  ROC_curvesCond <- c(ROC_curvesCond,list(PerCond$ROC_curve,PerCond$ROC_properties$Feature))
}



# Sum per ID
ds <- ds %>%
  group_by(ID,Cond) %>%
  mutate(SelfInter_persegmCond = sum(nLineCross)) 

for(Cond_i in 1:3) {
  PerCond <- ds %>%
    filter(Cond %in% Cond_list[Cond_i]) %>%
    
    Comp_ROC(., "group", "SelfInter_persegmCond","ID","moreCtl")
  PerCond$ROC_properties$Feature <- paste0(PerCond$ROC_properties$Feature, "_", Cond_list[Cond_i])
  All_ROC_Cond <- rbind(All_ROC_Cond, PerCond$ROC_properties)
  ROC_curvesCond <- c(ROC_curvesCond,list(PerCond$ROC_curve,PerCond$ROC_properties$Feature))
}


ggroc(list(ROC_curvesCond[[1]],ROC_curvesCond[[3]],ROC_curvesCond[[5]],ROC_curvesCond[[7]],ROC_curvesCond[[9]],ROC_curvesCond[[11]],ROC_curvesCond[[13]],ROC_curvesCond[[15]],ROC_curvesCond[[17]],ROC_curvesCond[[19]])) +
  scale_color_manual(labels = c(ROC_curvesCond[[2]],ROC_curvesCond[[4]],ROC_curvesCond[[6]],ROC_curvesCond[[8]],ROC_curvesCond[[10]],ROC_curvesCond[[12]],ROC_curvesCond[[14]],ROC_curvesCond[[16]],ROC_curvesCond[[18]],ROC_curvesCond[[20]]), values = brewer.pal(11, name = "Set3")) + 
  theme_minimal()


```

## Summary table Per cond

```{r}
knitr::kable(All_ROC_Cond[order(All_ROC_Cond$AUC, decreasing = TRUE), ])
```

# Graph

```{r}
ggplot(aes(x = GA_isValidStruct, y = group, fill = factor(stat(quantile))), data = ds_Q) +
  stat_density_ridges(
    geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 4, alpha = 0.5) +
  geom_vline(xintercept = All_ROC$threshold[All_ROC$Feature == "GA_isValidStruct"] , linetype = "dashed") +
  scale_fill_viridis_d(name = "Quartiles")


```

# References
